### 2017-06-01 writing/paper-Tian-gr-wetting.tex
Minor change and clean up files

% Created 2017-06-01 Thu 19:09
% Intended LaTeX compiler: pdflatex
\documentclass[journal=jacsat,manuscript=article,email=true]{achemso}
 \usepackage{minted}
 \usepackage{graphicx}
\usepackage{float}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{fontspec}
\keywords{Two-dimensional materials, doping, wettability, multiscale analysis, MD Simulation, eletrical double layer}
\author{Tian Tian}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\author{Siyu Li}
\affiliation{Key Laboratory of Energy Thermal Conversion and Control of Ministry of Education, School of Energy & Environment, Southeast University, Nanjing, Jiangsu 210096, China}
\author{Lingling Zhao}
\affiliation{Key Laboratory of Energy Thermal Conversion and Control of Ministry of Education, School of Energy & Environment, Southeast University, Nanjing, Jiangsu 210096, China}
\author{Elton J. G. Santos}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, United Kingdom}
\affiliation{School of Chemistry and Chemical Engineering, Queen's University Belfast, United Kingdom}
\author{Shangchao Lin}
\email{slin@eng.fsu.edu.}
\affiliation{Department of Mechanical Engineering, Materials Science and Engineering Program, FAMU-FSU College of Engineering, Florida State University, Tallahassee, Florida 32310, United States}
\author{Chih-Jen Shih}
\email{chih-jen.shih@chem.ethz.ch}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\date{}
\title{Multiscale Understanding the Wettability of Doped Two-Dimensional Materials}
\begin{document}

\newpage{}
\begin{abstract}
  Engineering molecular interactions at two-dimensional (2D) materials interfaces enables new technological opportunities 
  in functional surfaces and molecular epitaxy. Quantifying the interplay between the interfacial forces and electric 
  potential of 2D materials represents the crucial first step for understanding their wettability.
  Here we develop the first theoretical framework to model the wettability of the doped 2D 
  materials by properly bridging the multiscale physical phenomena at the 2D interfaces,
  including: (i) the change of 2D materials surface energy (atomistic scale), 
  (ii) the molecular reorientation of liquid molecules adjacent to the interface, 
  (molecular scale), and (iii) the electrical double layer (EDL) formed in the liquid phase (macroscopic scale). 
  The latter two effects are found to be the major mechanisms responsible for the contact angle change upon doping.
  When the doping level is electrostatically tuned, we demonstrate that 2D materials with high quantum capacitances 
  (e.g., transition metal dichalcogenides, TMDCs) possess a wider range of tunability in the interfacial tension. 
  Furthermore, practical considerations such as defect and airborne contamination are also quantitatively discussed. 
  Our analysis implies that the doping level can be another variable to modulate the wettability at the interface of 2D materials,
  as well as the molecular packing behavior on a 2D material-coated surface, 
  essentially faciliating the interfacial engineering of 2D materials.
\end{abstract}
# \maketitle

\newpage{}

\section{Introduction}
\label{sec:org435dbee}

Two-dimensional (2D) materials, the crystalline films with a thickness
of one atom or one molecule, combine optical transparency and the
quantum-confined properties
\cite{Novoselov_2005,Mas_Ballest__2011,Novoselov_2016}.  Recent progress in the scalable production of large-area 2D
materials \cite{Boott_2015,Li_2016} has paved the way for the
applications of functional surfaces and ultrathin membranes
\cite{Prasai_2012,Feng_2016,Rafiee_2010,Yin_2014,Surwade_2015,Hern_ndez_2013}.
The concept of van der Waals (vdW) epitaxy has been further extended
to grow molecular heterostructures on 2D material-coated surfaces for
a variety of applications
\cite{Shi_2012_epitaxy_graphene,Kim_2017_remote_epitaxy,Novoselov_2016}. The
realization of these technologies essentially requires precise control
over the molecular interactions at 2D material interfaces, which
motivates fundamental research of their interfacial properties, in
particular the wettability.  An accurate determination of the
interfacial tension, nevertheless, remains highly disputed
\cite{taherian2013what,Kozbial_2015,Parobek_2015,Govind_Rajan_2016}. Early
literature on understanding the surface science of 2D materials has
highlighted the effects of the airborne contaminants
\cite{li_effect_2013,Xu_2013_withwhat,kozbial_study_2014,Kozbial_2015,Chow_2015}
and the underlying substrate
\cite{raj_wettability_2013,rafiee_wetting_2012,shih_breakdown_2012,shih_wetting_2013},
since the length scale for the van der Waals (vdW) radius is
comparable to the thickness of a monolayer. On the other hand, a
subtle but important fact is that the 2D semimetals (e.g, graphene and
silicene) and 2D semiconductors (e.g., transition metal
dichalcogenides (TMDCs)) possess low density of states (DOSs) around
the intrinsic Fermi level (\(E_{\mathrm{F}}\)), such that the effect of
doping, either induced by the surroundings
\cite{Chen_2013,Varchon_2007,Giovannetti_2008}, or by the electrostatic
gating \cite{Das_2008,Perera_2013}, also comes into play. Recent advance
in the \textit{ab initio} calculations and scanning tunneling
microscopy (STM) studies on the adsorbed molecules on graphene
\cite{Muruganathan_2015,Huttmann_2015} has provided some clues to the
doping-dependent vdW interactions. Subsequently, the doping-induced
change in the wettability of graphene has been observed in the
contact-angle experiments
\cite{hong_mechanism_2016,goniszewski_correlation_2016,ashraf_doping-induced_2016},
and the molecular dynamics (MD) simulations
\cite{ostrowski_tunable_2014,ren_interfacial_2015,Taherian_2015,daub_electrowetting_2007}. In
this respect, however, most reports described the doping-dependent
wettability with the basic Young-Lippmann equation (YLE)
\cite{Lippmann_1875}, whereas the interplay between the orientation of
liquid molecules at the interface \cite{Shen_2006} and the electronic
structure of the 2D materials is often ignored. In order to address
the discrepancies among the literature, a general and complete
theoretical picture that bridges the gap between different length
scales, is clearly required.

In this report, we propose the first theoretical framework to model
the change of interfacial tension between liquid and a sheet of
monolayer 2D material. Multiscale physical phenomena are
considered. At the atomistic scale, we formulate the dependence of
surface energy of a 2D material on the doping density using the
quantum capacitance (QC) calculated by the density functional theory
(DFT). Next, the surface-charge-induced reorientation of liquid
molecules adjacent to the interface is associated with an N-body
system, which is resolved by molecular dynamics (MD) simulations,
allowing quantification of the interfacial tension in the absence of
electrolytes in liquid. The effect of the electrical double layer
(EDL) induced by the electrostatic interactions between the 2D
material surface charges and the ionic species in liquid, is then
addressed at the continuum level. Practical considerations, such as
the defect density and the surface contamination are also taken into
account to provide a comprehensive understanding of the
phenomena. Finally, we examine and validate our theory by comparing
with the contact angle changes reported in the electrowetting and the
substrate-induced doping experiments.

\section{Results and Discussions}
\label{sec:org66b0b42}

\subsection{Surface Energy of Doped 2D Materials}
\label{sec:org87cc68c}

Consider a liquid (\emph{L}) droplet sitting on a flat, monolayer 2D
material (\emph{2D}) supported by a solid substrate, following the Young’s
equation, the change of the equilibrium contact angle \(\theta\) upon
doping, is given by
\begin{equation}
\label{eqn-def-Young-Delta-theta}
\gamma_{\mathrm{L}} \Delta \cos\theta = \Delta \gamma_{\mathrm{2D}}
                                 - \Delta \gamma_{\mathrm{2D-L}}
\end{equation}
where \(\gamma_{\mathrm{L}}\) and \(\gamma_{\mathrm{2D}}\) are the surface
tensions of liquid and the 2D material considered,
respectively, \(\gamma_{\mathrm{2D-L}}\) is the interfacial tension between the
liquid and the contacting 2D material. We define \(\Delta \gamma =
\gamma - \gamma_{0}\) for all the surface (interfacial) tensions considered and
\(\Delta \cos \theta = \cos \theta - \cos \theta_{0}\), where the
subscript 0 corresponds to that in the case of intrinsic 2D material,
with the doping density per unit area, \(\sigma_{\mathrm{2D}} = 0\). In
the theoretical analysis presented here, we aim to model \(\Delta \cos
\theta\) as a function of \(\sigma_{\mathrm{2D}}\). Note that under the
assumptions that (i) the change of \(E_{\mathrm{F}}\) in the doped 2D
materials does not result in the interfacial electron transfer,
namely, the electrochemical reactions, (ii) the doping effect does not
change the surface energy of the underlying substrate, and (iii) the
vdW and electrostatic interactions are perfectly additive and
pairwise, one can decouple the effect of the underlying
substrate. Therefore the debate of the “wetting transparency”
\cite{rafiee_wetting_2012,shih_wetting_2013}, does not affect our analysis.

First, we model the dependence of \(\gamma_{\mathrm{2D}}\) on the doping
level. Consider a closed system with constant pressure and volume
containing a sheet of free-standing 2D material, following the Euler
homogeneous function theorem of thermodynamics, the total internal
energy of the system \(U\), is given by \(U = TS + \mu_{\mathrm{2D}} N +
\gamma_{\mathrm{2D}} A + \psi_{\mathrm{2D}} q\)
\cite{bard_electrochemical_1980}, where \(T\) is the temperature, \(S\) is
the entropy, \(\mu_{\mathrm{2D}}\) is the chemical potential of the 2D
material per unit lattice, \(N\) is the number of unit lattices, \(A\) is
the area of the 2D material, and \(\psi_{\mathrm{2D}}\) and \(q\) are the
electric potential and total charge in the 2D material,
respectively. At constant \(T\), combining with the first law of
thermodynamics and the differential form of \(U\), it follows:
\begin{equation}
\label{eqn-dgamma-dpsi}
\mathrm{d} \gamma_{\mathrm{2D}} = -\frac{q}{A} \mathrm{d} \psi_{\mathrm{2D}}
                                = -\sigma_{\mathrm{2D}} \mathrm{d} \psi_{\mathrm{2D}}
\end{equation}
relating the surface tension change of a free-standing 2D material as
a function of its doping density \(\sigma_{\mathrm{2D}}\). After bringing an amount of charge
\(q\) from infinity to a charge-neutral 2D material in the
aforementioned system, the surface energy change is therefore given
by:
\begin{equation}
\label{eqn-delta-gamma-sigma-free-2D}
\Delta \gamma_{\mathrm{2D}} = - \int_{0}^{\psi_{\mathrm{2D}}} \sigma_{\mathrm{2D}} \mathrm{d}\psi'
                            = - \int_{0}^{\sigma_{\mathrm{2D}}} \sigma' \left( \frac{1}{C_{\mathrm{2D}}}\right) \mathrm{d} \sigma'
\end{equation}
where \(C_{\mathrm{2D}} = g(E_{\mathrm{F}}) e^{2}\) is the quantum
capacitance of the 2D material \cite{davies_two-dimensional_1997,Das_Sarma_2011},
\(g(E_{\mathrm{F}})\) is the DOS as a function of \(E_{\mathrm{F}}\),
\(\psi_{\mathrm{2D}} = -(E_{\mathrm{F}} - E_{\mathrm{F,0}})/e\) and
\(E_{\mathrm{F,0}}\) corresponds to the Fermi level of the 2D material
at the charge neutral point (CNP). Accordingly, eq
\ref{eqn-delta-gamma-sigma-free-2D} provides a simple relation which
depicts the surface tension change of a 2D material at the
quantum-mechanical level. We have calculated the DOSs as a function of
\(E_{\mathrm{F}}\) for a variety of 2D materials using the density
functional theory \cite{tian_multiscale_2016}. Note that we take into
account a fractional component of the exact exchange from the
Hartree-Fock (HF) theory hybridized with the DFT exchange-correlation
functional at the level of the HSE06 hybrid functional. Therefore, any
limitation of the exchange and correlation functional utilized in the
chemical description of the energy levels can be improved. The
following 2D materials are considered: TMDC monolayers (MX\(_{\text{2}}\), M =
Mo, W and X = S, Se, Te), silicene, germanene, phosphorene (monolayer
black phosphorus), and graphene. The doping density in a 2D material
is calculated by integrating the DOS from its intrinsic Fermi level,
i.e., \(\sigma_{\mathrm{2D}} = \int_{E_{\mathrm{F,0}}}^{E_{\mathrm{F}}}
g(E')e \mathrm{d}E'\) \cite{john_quantum_2004}.

Figure \ref{fig:dgamma-sigma} presents the calculated \(\Delta
\gamma_{\mathrm{2D}}\) as a function of \(\sigma_{\mathrm{2D}}\) for the
2D materials considered here. Clearly, the doping of 2D materials
reduces their surface energy, or based on the classical definition,
the work required to separate two stacked monolayers is lowered. Among
the 2D materials, we find that graphene shows the highest degree of
surface energy decrease, up to -16 mJ\(\cdot \mathrm{m}^{-2}\) , or
\textasciitilde{}20\% reduction of its intrinsic surface tension
\cite{shih_wetting_2013}, at \(\sigma_{\mathrm{2D}}\) = \(\pm
4\times10^{13}\ e\cdot \mathrm{cm}^{-2}\). However, as will be
discussed later, it does not imply a reduced wettability because the
quantum capacitance effect also reduces the interfacial tension,
\(\gamma_{\mathrm{2D-L}}\). A clear trend is that the surface energy
decrease is more significant in the 2D semimetals (e.g. graphene,
silicene, and germanene) than that in the 2D semiconductors
(e.g. TMDCs). This reflects the fact that the effective mass of
carriers in the 2D semiconductors is much higher than that in the 2D
semimetals \cite{davies_two-dimensional_1997}, thereby resulting in
high DOS, as well as a high \(C_{\mathrm{2D}}\) (see eq
\ref{eqn-delta-gamma-sigma-free-2D}). This concept also explains why the
surface energy decrease for silicene and germanene are lower than
that for graphene \cite{Yan_2013}.  To our knowledge, the doping-induced
surface energy change in 2D materials has never been investigated
experimentally, which may be of interest for future study.

\subsection{Reorientation of Liquid Molecules}
\label{sec:org188528c}

Next, we discuss the interactions between 2D materials and liquid. In
a doped 2D material, the delocalized carriers are confined in the 2D
plane. Therefore, following the spirit of the mean-field theory, we
treat it as a continuously, uniformly charged surface. Since these
charges are either generated by interacting with the underlying
substrate, or electrostatically induced by gating, the
electroneutrality still holds before in contact with liquid. The
surface charges result in two consequences that may change
\(\gamma_{\mathrm{2D-L}}\), including (i) the reorientation of adjacent
liquid molecules \cite{ostrowski_tunable_2014} and (ii) the formation of
the electric double layer (EDL) at the liquid-solid interface, known
as the electrowetting effect
\cite{Lippmann_1908,mugele_electrowetting:_2005}. Under the assumption
that the vdW and electrostatic (Coulombic) interactions are additive,
which compose the interfacial tension, we propose that the
interfacial tension change is given by:

\begin{equation}
\label{eqn-delta-gamma-decompose}
\Delta \gamma_{\mathrm{2D-L}} = \Delta \gamma_{\mathrm{2D-L}}^{\mathrm{Orien}}
                              + \Delta \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}}
\end{equation}
where \(\Delta \gamma_{\mathrm{2D-L}}^{\mathrm{Orien}}\) and \(\Delta
\gamma_{\mathrm{2D-L}}^{\mathrm{EDL}}\) correspond to the contributions
from the reorientation and the EDL effects, respectively, as
schematically illustrated in Figure \ref{fig:scheme-method}. We also
assume that \(\Delta \gamma_{\mathrm{2D-L}}^{Orien}\) is independent of
the electrolyte concentration, since the concentration of liquid
molecules is typically orders-of-magnitude higher.

First we discuss the orientation effect. Understanding the
reorientation effect involves positioning and sampling the collective,
time-averaged motion of liquid molecules near the interface, which is
a standard molecular dynamics (MD) problem. Note that even with the
state-of-the-art MD algorithms, it remains challenging to accommodate
the calculations for the EDL, in which the length scale of electric
field can be larger than one micrometer in diluted electrolyte
solutions, e.g., pure water with self-ionized H\(_{\text{3}}\)O\(^{\text{+}}\) and OH\(^{\text{-}}\)
ions. Here we consider the graphene-water interface as a model
system. All MD simulations were carried out using the GROMACS 4.5
software package \cite{Hess_2008}. Monolayer graphene was modeled as an
infinite rigid sheet in the x-y plane. The carbon atoms of graphene
were treated as uncharged Lennard-Jones (LJ) spheres with \(\sigma\) =
0.34 nm and \(\epsilon\) = 0.223 kJ/mol \cite{Cheng_1990}, using the
force-field parameters reported by Tummala and Striolo
\cite{Tummala_2008}. The doping effect is included by assigning an equal
amount of charge \(\sigma_{\mathrm{2D}} / \rho_{\mathrm{G}}\), where
\(\rho_{\mathrm{G}}\) is the surface density of carbon atoms, to each
carbon atom. The \(\sigma_{\mathrm{2D}}\) range considered here
approximately correspond to the partial atomic charge from −0.012 to
0.012 \(e/\mathrm{atom}\). Water molecules were modeled using the SPC/E
model \cite{Berendsen_1987} with bond lengths and angles of water
molecules constrained using the SETTLE algorithm
\cite{Miyamoto_1992}. Lennard-Jones interactions were treated with a
cutoff distance of 1 nm, with those between different atoms calculated
using the standard geometric averaging rule. Long-range electrostatic
interactions were treated using the particle mesh Ewald (PME)
summation method \cite{Darden_1993,Essmann_1995} with a short-range
cutoff distance of 1 nm. The velocity-rescaled Berendsen thermostat
was implemented to maintain a constant system temperature of 298.15 K
\cite{Bussi_2007}. All simulations were carried out under the NVT
ensemble.

In order to precisely determine the interfacial interactions using MD
simulations, instead of the commonly used model that compared the
nanoscale contact angle by placing a nanodroplet onto a sheet of
suspended 2D material
\cite{ostrowski_tunable_2014,daub_electrowetting_2007,ren_interfacial_2015,Taherian_2015},
we simulate the difference of the total potential energy, \(E\), between
two separate systems that contain (i) only water molecules with two
surfaces exposing to vacuum (L), and (ii) the same amount of water
molecules with one surface in contact with graphene (placed at z = 0)
and the other surface exposing to vacuum (GL) (see Figure
\ref{fig:MD-res}(a) and \ref{fig:MD-res}(b)) . Periodic boundary
conditions are used in all three directions of the simulation boxes in
both systems. Additionally, a vacuum layer of 3 nm thick along the
\emph{z}-axis is placed to separate the periodic images of the
graphene-water system. The total energy in both systems can be
formulated as: \(E_{\mathrm{L}} = \mu_{\mathrm{L}}n_{\mathrm{L}} +
2\gamma_{\mathrm{L}}S\) and
\(E_{\mathrm{GL}}=\mu_{\mathrm{L}}n_{\mathrm{L}}+(\gamma_{\mathrm{L}} +
\gamma_{\mathrm{2D-L}} + \gamma_{\mathrm{2D}})S\), respectively, where
\(\mu_{\mathrm{L}}\) is the chemical potential per water molecule in the
bulk phase, \(n_{\mathrm{L}}\) is the number of liquid molecules in the
simulation box, and \(S\) is the area of the \emph{xy}-plane.  A simulation
box with a sufficiently large length of 21 nm in the \emph{z}-direction is used,
with a 18-nm thick block of water molecules, to minimize the effect of
the long-range electrostatic interaction between the charged graphene
and the water molecules at the water-vacuum interface, by ensuring
that the time-averaged dipole moment for the water molecules at the
water-vacuum interface is approximately zero (see Supporting Information
Figure S1).  One can show that \(E_{\mathrm{GL}} - E_{\mathrm{L}} =
(\gamma_{\mathrm{2D-L}} + \gamma_{\mathrm{2D}} - \gamma_{L})S =
(\Phi + 2 \gamma_{\mathrm{2D}})S\), where \(\Phi\) is the interfacial
energy, which is defined as \(\Phi = \gamma_{\mathrm{2D-L}} -
\gamma_{\mathrm{2D}} - \gamma_{\mathrm{L}}\), and combining with eq
\ref{eqn-def-Young-Delta-theta}, the change of interfacial energy
\(\Delta \Phi\) can be formulated as:
\begin{equation}
\label{eqn-Delta-Phi-Delta-cos}
\Delta \Phi = \Delta (E_{\mathrm{GL}} - E_{\mathrm{L}})/S - 2\Delta \gamma_{\mathrm{2D}} = -\gamma_{\mathrm{L}} (\Delta \cos \theta)^{\mathrm{Orien}}
\end{equation}
where \((\Delta\cos \theta)^{\mathrm{Orien}}\) corresponds to the
contact angle change due to the reorientation effect. Note that in the
current MD simulation setup, the surface tension of the 2D material
remains unchanged (i.e. \(\Delta \gamma_{\mathrm{2D}}=0\)), since the
Fermi level change with
respect to \(\sigma_{\mathrm{2D}}\) (see eq
\ref{eqn-delta-gamma-sigma-free-2D}) is not included in the algorithm. In all simulations we assume a
constant surface tension of water \(\gamma_{\mathrm{L}}=72.8\
\mathrm{mJ}\cdot \mathrm{m}^{-2}\). Accordingly, \(\Delta \Phi\) follows
\(\Delta \Phi = \Delta (E_{\mathrm{GL}} - E_{\mathrm{L}})/S = \Delta
\Phi_{\mathrm{LJ}} + \Delta \Phi_{\mathrm{Coul}}\), where \(\Delta
\Phi_{\mathrm{LJ}}\) and \(\Delta \Phi_{\mathrm{Coul}}\) correspond to
the change for the vdW and Coulombic interaction potentials,
respectively.

Figure \ref{fig:MD-res}(c) shows the values of \(\Delta \Phi\), and its
contributions from the Lennard-Jones (\(\Delta \Phi_{\mathrm{LJ}}\)) and
Coulombic interactions (\(\Delta \Phi_{\mathrm{Coul}}\)), as functions
of \(\sigma_{\mathrm{2D}}\). We note that the change of adhesion energy
in the doped graphene system is dominated by the Coulombic
interaction. When the doping level of graphene is \(\pm 4 \times
10^{13}\ e\cdot \mathrm{cm}^{-2}\), the Coulombic interaction causes a
decrease in the adhesion energy of -10\textasciitilde{}-15 mJ\(\cdot \mathrm{m}^{-2}\),
while the vdW interaction, on the other hand, causes a slight increase
in the adhesion energy by less than 5 mJ\(\cdot \mathrm{m}^{-2}\). In
other words, concerning the reorientation of water molecules at high
doping levels, the Coulombic interaction favors the decrease of
\(\gamma_{\mathrm{2D-L}}\), and thus the contact angle \(\theta\), while
the vdW interaction slightly increases \(\gamma_{\mathrm{2D-L}}\) and
\(\theta\). We shall note that due to the highly polar nature of water
molecules, the average inter-molecular equilibrium distance in the
absence of external electric field is shorter than the Lennard-Jones
equilibrium distance. Increasing the doping density of the graphene
sheet essentially enhances the electrostatic attractions between
graphene and water molecules and further decreases the inter-molecular
equilibrium distance, as reflected from the red shifts of the first
peaks in Figure \ref{fig:MD-res}(d) for the two charged graphene cases
compared to the charge-neutral case, which eventually leads to a
slight increase in .  The plot of \(\Delta \Phi\) as a function of
\(\sigma_{\mathrm{2D}}\) in Figure \ref{fig:MD-res}(b) shows an apparent
asymmetric shape. This is can be further revealed from the
\emph{z}-dependent local molecular density \(\rho_{\mathrm{L}}\) and charge
density \(\delta_{\mathrm{L}}\) profiles of the water molecules, as
shown in Figure \ref{fig:MD-res}(d) and \ref{fig:MD-res}(c),
respectively. The molecular density of the first water layer adjacent
to graphene increases when graphene is p-doped
(\(\sigma_{\mathrm{2D}}=0.012\ e\cdot \mathrm{cm}^{-2}\), or
equivalently 4.56 \textit{e}/atom) and decreases when graphene is
n-doped (\(\sigma_{\mathrm{2D}}=-0.012\ e\cdot \mathrm{cm}^{-2}\), or
equivalently 4.56 \textit{e}/atom) compared with the case of
charge-neutral graphene, indicating the polarity-dependent adsorption
of water molecules on graphene. Similar trend can also be observed in
the \(\delta_{\mathrm{L}}\) profiles. Interestingly, \(\Delta \Phi\)
exhibits positive Coulombic contribution for slightly n-doped
graphene, when \(\sigma_{\mathrm{2D}} \approx -2\times10^{13}\ e\cdot
\mathrm{cm}^{-2}\), indicating that the reorientation of interfacial
water molecules by small negative charges on graphene helps minimizing
the Coulombic interactions.
\begin{center}
\fbox{
\begin{minipage}[c]{.8\linewidth}
\textbf{\textsf{\textsc{TODO}}} Explanation of the positive part of \(\Delta \Phi\) when \(\sigma_{\mathrm{2D}} \approx -2\) ?

\end{minipage}
}
\end{center}
Although the MD simulation results show the Coulombic interaction
between the doped graphene and water molecules dominate the
reorientation effect, such conclusion may not be readily applied to
other types of 2D materials. For 2D semimetal systems (silicene and
germanene), the intensity of Coulombic interactions may vary due to
the mismatch among the lattice constants. On the other hand, for the
2D semiconductors such as the TMDCs, in which their
\(\gamma_{\mathrm{2D-L}}\) are already largely determined by the
Coulombic interactions in the intrinsic case
\cite{Govind_Rajan_2016,Chow_2015}. Therefore we anticipate that
case-by-case studies and DFT calculations which deal with the
post-doping charge redistribution, might be required.
\begin{center}
\fbox{
\begin{minipage}[c]{.8\linewidth}
\textbf{\textsf{\textsc{TODO}}} Try find other references?

\end{minipage}
}
\end{center}

\subsection{Electrical Double Layer Effect}
\label{sec:org0491aea}

The EDL effect that decreases the interfacial tension has been
extensively studied in the context of the electrowetting phenomena
\cite{Mugele_2005}, in which the interfacial tension is reduced due to
the adsorption of ionic species at the solid-liquid interface. Here we
extend the concept to the 2D material-liquid systems. Consider a
closed system containing an interface formed between liquid and a
sheet of free-standing 2D material, by combining with the Gibbs
adsorption theory \cite{bard_electrochemical_1980}, one can obtain:
\begin{equation}
\label{eq:2}
\mathrm{d} \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}} = -\sigma_{\mathrm{2D}} \mathrm{d} \psi_{\mathrm{2D}}
                                                   -\sum_{\mathrm{i}} \Gamma_{\mathrm{i}} \mathrm{d}\mu_{\mathrm{i}}
\end{equation}
where \(\Gamma_{\mathrm{i}}\) and \(\mu_{\mathrm{i}}\) are the interfacial
excess and the chemical potential of the ionic specie \(i\) at the
interface, respectively. Relative to that in the bulk phase
(\(\psi=0\)), the interfacial chemical potential of a charged solute
\(\mu_{\mathrm{i}}\), is given by
\(\mu_{\mathrm{i}}=z_{\mathrm{i}}e\psi_{\mathrm{2D}}\), where
\(z_{\mathrm{i}}\) is the valency of component \(i\), under the assumption
that the interfacial electric potential is equal to
\(\psi_{\mathrm{2D}}\). As with the approach leading to eq
\ref{eqn-delta-gamma-sigma-free-2D}, it follows:
\begin{equation}
\label{eqn-Delta-gamma-2DL-EDL-full}
\Delta \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}} = 
-\int_{0}^{\sigma_{\mathrm{2D}}} \sigma' \left(\frac{1}{C_{\mathrm{2D}}}\right) \mathrm{d}\sigma'
-\int_{0}^{\sigma_{\mathrm{L}}} \sigma' \left(\frac{1}{C_{\mathrm{EDL}}}\right) \mathrm{d}\sigma'
\end{equation}
where \(\sigma_{\mathrm{L}}\) is the interfacial charge density per unit
area in the liquid phase, and \(C_{\mathrm{EDL}}\) is the EDL
capacitance. Combining eq \ref{eqn-Delta-gamma-2DL-EDL-full} with eqs
\ref{eqn-def-Young-Delta-theta} and \ref{eqn-delta-gamma-sigma-free-2D},
the contact angle change due to the EDL effect, \((\Delta \cos
\theta)^{\mathrm{EDL}}\) is given by:
\begin{equation}
\label{eqn-Delta-cos-EDL}
(\Delta \cos \theta)^{\mathrm{EDL}} = -\frac{1}{\gamma_{\mathrm{L}}}
                                      \int_{0}^{\sigma_{\mathrm{L}}} \sigma' 
                                      \left(\frac{1}{C_{\mathrm{EDL}}}\right) \mathrm{d}\sigma'
\end{equation}
In other words, eqs \ref{eqn-delta-gamma-sigma-free-2D} and
\ref{eqn-Delta-gamma-2DL-EDL-full} suggest that \(\sigma_{\mathrm{2D}}\)
decreases both \(\gamma_{\mathrm{2D}}\) and \(\gamma_{\mathrm{2D-L}}\) to
the same extent, i.e. \(\int_{0}^{\sigma_{\mathrm{2D}}}
\sigma'/C_{\mathrm{2D}} \mathrm{d}\sigma'\), such that the contribution
of quantum capacitance cancels out in the YLE.

For a \(z:z\) electrolyte solution, with the molar concentration \(c_{0}\)
and the electrolyte valence \(z\), we describe the electric potential
profile in liquid with the one-dimensional Gouy-Chapman-Stern model
\cite{Stern_1924_theory}, as schematically shown in Figure
\ref{fig:res-EDL}(a).  The EDL consists of a Helmholtz layer with the
permittivity of liquid \(\epsilon_{\mathrm{L}}\) and thickness
\(d_{\mathrm{H}}\), next to the Gouy-Chapman (GC) diffuse layer where
the electric potential profile follows the Poisson Boltzmann equation
\cite{bard_electrochemical_1980}. The \(C_{\mathrm{EDL}}\) of the
Gouy-Chapman-Stern model is equivalent to two capacitors in series,
namely,
\begin{equation}
\label{eqn-GCS-capacitance-series}
\frac{1}{C_{\mathrm{EDL}}} = \frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}
\end{equation}
where \(C_{\mathrm{H}}=\epsilon_{\mathrm{L}}/d_{\mathrm{H}}\) is the
capacitance of the Helmholtz double layer and
\(C_{\mathrm{GC}}=\sqrt{\frac{2z^{2}e^{2}\epsilon_{\mathrm{L}}c_{0}N_{\mathrm{A}}}{k_{\mathrm{B}}T}}
\cosh(\frac{ze\psi_{\mathrm{GC}}}{2k_{\mathrm{B}}T})\) is the
differential capacitance of the GC layer, with \(\psi_{\mathrm{GC}}\)
corresponding to the electric potential at the GC interface. Consider
a Gaussian enclosure including the 2D material layer and the EDL, due
to a zero net electric field at the surface of the Gaussian enclosure,
the electroneutrality of the system \cite{bard_electrochemical_1980}
suggests:
\begin{equation}
\label{eqn-neutrality}
\sigma_{\mathrm{2D}} + \sigma_{\mathrm{L}} = 0
\end{equation}
And therefore, the electric potentials corresponding to the surface of
the 2D material and the outer Helmholtz plane, \(\psi_{\mathrm{2D}}\)
and \(\psi_{\mathrm{GC}}\), are given by:
\begin{equation}
\label{eqn-psi-GC}
\begin{aligned}
  \psi_{\mathrm{GC}} &= -\frac{2k_{\mathrm{B}}T}{ze} 
                         \sinh^{-1}\left(
                         \frac{-\sigma_{\mathrm{2D}}}{\sqrt{8c_{0}N_{\mathrm{A}}\epsilon_{\mathrm{L}}k_{\mathrm{B}}T}}
                            \right) \\
  \psi_{\mathrm{2D}} &= \psi_{\mathrm{GC}} + \sigma_{\mathrm{2D}}\frac{d_{\mathrm{H}}}{\epsilon_{\mathrm{L}}}
\end{aligned}
\end{equation}
where \(N_{\mathrm{A}}\) is the Avogadro constant and \(k_{\mathrm{B}}\)
is the Boltzmann constant. Accordingly, the final form
of \((\Delta \cos \theta)^{\mathrm{EDL}}\) is given by:
\begin{equation}
\label{eqn-Delta-cos-EDL-final}
\begin{aligned}
(\Delta \cos \theta)^{\mathrm{EDL}} &= \frac{\sigma_{\mathrm{2D}}^{2}}{2\gamma_{\mathrm{L}}
                                       C_{\mathrm{H}}}\frac{1}{\gamma_{\mathrm{L}}}
                                       \sqrt{\frac{32k_{\mathrm{B}}^{3}T^{3} \epsilon_{\mathrm{L}} c_{0} 
                                       N_{\mathrm{A}}}{z^{2}e^{2}}} 
                                       \left[\cosh(\frac{z e \psi_{\mathrm{GC}}}{2k_{\mathrm{B}} T}) -1 \right]\\
                                    &= \frac{\Delta \sigma_{\mathrm{2D}}^{2}}{2\gamma_{\mathrm{L}}}
                                       \left[ \frac{1}{C_{\mathrm{H}}} + \frac{1}{(C_{\mathrm{DH}} + 
                                       C_{\mathrm{GC}})/2}\right]
\end{aligned}
\end{equation}
where \(C_{\mathrm{DH}}=\epsilon_{\mathrm{L}}/\lambda_{\mathrm{D}}\) is
the differential capacitance of the diffuse layer using the
Debye-Hückel approximation \cite{bard_electrochemical_1980}, and
\(\lambda_{\mathrm{D}}=\sqrt{\frac{\epsilon_{\mathrm{L}}k_{\mathrm{B}}T}{2z^{2}e^{2}c_{0}N_{\mathrm{A}}}}\)
is the Debye screening length. Compared with the classical YLE,
\(\Delta \cos \theta =
-\frac{\sigma^{2}}{2\gamma_{\mathrm{L}}C_{\mathrm{YL}}}\), where
\(C_{\mathrm{YL}}\) is the equivalent capacitance in YLE, one can show that \(C_{\mathrm{YL}}\)
is given by:
\begin{equation}
\label{eqn-CYL-model}
\frac{1}{C_{\mathrm{YL}}} = \frac{1}{C_{\mathrm{H}}} + \frac{1}{(C_{\mathrm{DH}} + C_{\mathrm{GC}})/2}
\end{equation}

The above analysis leads to two important observations: (i) If a thin
contamination layer (usually composed of charge-neutral hydrocarbon
species \cite{li_effect_2013}) exists between the Helmholtz layer and
the 2D material surface, the electroneutrality at the interface (eq
\ref{eqn-neutrality}) still holds, such that \((\Delta \cos
\theta)^{\mathrm{EDL}}\) remains the same (see eq
\ref{eqn-Delta-cos-EDL-final}). In other words, despite the recent
debate about the role of the airborne adsorbates in the contact angle
measurement on 2D materials
\cite{li_effect_2013,Kozbial_2015,Xu_2013_withwhat,Chow_2015}, we expect
that the experimental characterization of the doping-induced contact
angle change is more robust and reproducible. (ii) As shown in eqs
\ref{eqn-GCS-capacitance-series} and \ref{eqn-Delta-gamma-2DL-EDL-full},
the equivalent capacitance in the YLE, \(C_{\mathrm{YL}}\), is not
identical to the total capacitance in the EDL, \(C_{\mathrm{EDL}}\). We
notice that the literature in this field (e.g. Refs.
\citenum{ostrowski_tunable_2014,daub_electrowetting_2007,goniszewski_correlation_2016,ashraf_doping-induced_2016})
often simply assumed that \(C_{\mathrm{YL}}=C_{\mathrm{DH}}\) in the
YLE, which significantly overestimates \((\Delta cos
\theta)^{\mathrm{EDL}}\), when \(\psi_{\mathrm{2D}} \gg
k_{\mathrm{B}}T/e\) \cite{israelachvili03_inter}, as can be found in heavily-doped 2D materials.

Consider the interface formed between graphene and an (1:1) aqueous
solution, using the parameters of \(d_{\mathrm{H}}=3\) Å
\cite{mcclendon_thickness_1927}, and \(\gamma_{\mathrm{L}}=72.8\
\mathrm{mJ}\cdot \mathrm{m}^{-2}\) at \(T=298\) K, the calculated
\((\Delta \cos \theta)^{\mathrm{EDL}}\) as a function of
\(\sigma_{\mathrm{2D}}\) is shown in Figure \ref{fig:res-EDL}(b).  Note
that \(c_{0}=10^{-7}\) mol\(\cdot \mathrm{L}^{-1}\) corresponds to the
pure water system. We find that the contact angle change is
stronger in a more diluted electrolyte solution, resulting from an
increase in Debye screening length \(\lambda_{\mathrm{D}}\), which
in turn decreases both \(C_{\mathrm{GC}}\) and \(C_{\mathrm{DH}}\). This
is distinct from the that in the typical electrowetting experiment on
a dielectric layer, in which \(C_{\mathrm{YL}}\) is governed by the
capacitance of the dielectric layer, so the contact angle change is
almost independent of the electrolyte concentration
\cite{Mugele_2005}. Moreover, within the range of \(\sigma_{\mathrm{2D}}\)
considered here (\(-4\times10^{13}\) \textasciitilde{} \(4\times10^{13}\ e\cdot
\mathrm{cm}^{-2}\)), the calculated \((\Delta \cos
\theta)^{\mathrm{EDL}}\) is always less than 0.06.  In combination with
the \(\Delta \gamma^{\mathrm{Orien}}_{\mathrm{2D-L}}\) (equivalent to
\(\Delta \Phi\), see previous section) calculated by the MD simulations,
the total change of contact angle \((\Delta \cos
\theta)^{\mathrm{Orien + EDL}}\) due to the doping effect is shown in
Figure \ref{fig:res-EDL}(c). The MD-calculated \(\Delta \Phi\) as
a function of \(\sigma_{\mathrm{2D}}\) is fitted by a third-order
polynomial function (see Supporting Information Figure S3).  We find
that the reorientation effect is more predominant than the EDL effect,
therefore the overall change of contact angle is less dependent on the
solute concentration, compared with Figure \ref{fig:res-EDL}(b).

Together with the reorientation effect, the calculated \(\Delta \cos
\theta\) change only reaches 0.06\textasciitilde{}0.08, corresponding to a maximum
contact angle decrease of 3.5\textasciitilde{}4.5\(^{\circ}\), for
\(\sigma_{\mathrm{2D}}=\pm1\times10^{13}\ e\cdot \mathrm{cm}^{-2}\),

which is is a typical range for both the electrostatic gating
\cite{hong_mechanism_2016} and substrate-induced doping
\cite{ashraf_doping-induced_2016}. Clearly, the predicted contact angle
change upon doping is still lower than those observed
experimentally. In the next section, we discuss the practical
considerations that may influence a 2D material-coated surface.

\subsection{Practical Considerations}
\label{sec:org80c83d8}

In practice, in order to
carry out the contact angle measurement, a large-area sheet of 2D
material is required to cover the surface. Therefore, most of the
experimental findings are based on the chemical-vapor-deposited (CVD)
samples, which are inherently polycrystalline with grain boundaries
and point defects \cite{Banhart_2011_defect}.  In addition, the transfer
of the ultrathin films onto a solid surface often results in
nanometer- to micrometer- scale tears and wrinkles. Although our
analysis has suggested that the charge-neutral polymer residues and
airborne contaminations do not influence \((\Delta
\theta)^{\mathrm{EDL}}\), here we show that the defect density plays an
important role in determining the wettability of doped 2D
materials.

Figures \ref{fig:f-nc-exp}(a) and \ref{fig:f-nc-exp}(b) illustrate how an
 electric field, either from the dopants on the substrate surface or
 from the electrostatic gating, penetrate through a void in a 2D
 material sheet and interact with the liquid phase directly. As a
 result, the EDL is built up adjacent to the substrate surface, with
 the surface excess \(\sigma_{\mathrm{v}}\) and the effective
 capacitance \(C_{\mathrm{v}}\).  Since \(\sigma_{v}\) is equivalent to
 \(\sigma_{\mathrm{2D}}\), together with the EDL and the reorientation
 effects discussed earlier, the modified YLE considering a 2D material
 with voids follows:
\begin{equation}
\label{eqn-def-Delta-cos-mixture}
\Delta \cos \theta = -f\frac{\sigma_{\mathrm{2D}}^{2}}{\gamma_{\mathrm{L}} C_{\mathrm{v}}} 
                     + (1-f)[(\Delta \cos \theta)^{\mathrm{Orien}} + (\Delta \cos \theta)^{\mathrm{EDL}}]
\end{equation}
where \(f\) is the void (defect) fraction in the 2D material. For the
electrostatic gating experiments (see Figure
\ref{fig:f-nc-exp}(a)), the 2D material quantum capacitor
and the dielectric  capacitor are connected in
series, so that the voltage applied between the gate electrode and
2D material, \(V_{\mathrm{G}}\), is given by \cite{tian_multiscale_2016}:
\begin{equation}
\label{eqn-VG-gating}
V_{\mathrm{G}} = \frac{\sigma_{\mathrm{2D}} - \sigma_{\mathrm{0}}}{C_{\mathrm{d}}}
                  + \int_{\sigma_{0}}^{\sigma_{\mathrm{2D}}} \frac{1}{C_{\mathrm{2D}}} \mathrm{d}\sigma'
\end{equation}
where \(C_{\mathrm{d}}\) is the capacitance of the dielectric layer, and
\(\sigma_{0}\) is the initial doping density of the 2D material,
corresponding to \(V_{\mathrm{G}}=0\).

Next, in order to examine the effect of incomplete 2D material
coverage, two independent sets of experimental results, which measure
the water contact angle on (i) substrate-doped graphene
\cite{ashraf_doping-induced_2016} and (ii) electrostatically-gated
graphene \cite{hong_mechanism_2016} are chosen to compare, with
\(C_{\mathrm{v}} = C_{\mathrm{DH}}\) and \(C_{\mathrm{v}} =
C_{\mathrm{d}}\) in eq \ref{eqn-def-Delta-cos-mixture}, respectively. The
parameter \(f\) and \(\sigma_{0}\) are determined by least-square fitting
the experimentally-observed \(\Delta \cos \theta\) with respect to
\(\sigma_{\mathrm{2D}}\) using eq \ref{eqn-def-Delta-cos-mixture}. Figure
\ref{fig:f-nc-exp}(c) compares the calculated \(\Delta \cos \theta\) as a
function of \(\sigma_{\mathrm{2D}}\), together with the two sets of
experimental data. In both cases considered, we observe a slight
degree of shift in the minima of the fitted curves, corresponding to
the CNP of graphene, or \(\sigma_{\mathrm{2D}} = 0\). In other words, we
observe \(\sigma_{0} \neq 0\) in both cases, which is typical for
CVD-grown graphene samples
\cite{Shih2015PartiallyScreened,goniszewski_correlation_2016}.  The
fitted values of \(f\) are reasonably small (3.6\% for the
electrostatically-gated graphene and 3.0\% for substrate-doped
graphene), clearly demonstrating that the contact angle change can be
greatly influenced by the defect density. We believe this explains the
discrepancy between the experimental observations and the multiscale
theoretical framework proposed here.

Finally, we discuss the influence of 2D material choice under
electrostatic gating condition. The above analysis (eq
\ref{eqn-Delta-cos-EDL-final}) has clearly suggested that the contact
angle change effect \((\Delta \cos \theta)^{\mathrm{EDL}}\) only depends
on \(\sigma_{\mathrm{2D}}\), which can be controlled by an electric
displacement field, with the experimental setup shown in Figure
\ref{fig:dcos-all-2D}(a). Specifically,  eqs \ref{eqn-psi-GC} and
\ref{eqn-VG-gating} suggest:
\begin{equation}
\label{eqn-dVG-choice-2D}
\begin{aligned}
\mathrm{d} V_{\mathrm{G}} &= (\frac{1}{C_{\mathrm{d}}} + \frac{1}{C_{\mathrm{2D}}}) \mathrm{d} \sigma_{\mathrm{2D}} \\
\mathrm{d} \psi_{\mathrm{2D}} &= -(\frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}) \mathrm{d} \sigma_{\mathrm{L}}
\end{aligned}
\end{equation}
And since \(\sigma_{\mathrm{2D}} = -\sigma_{\mathrm{L}}\), the first
derivative of \(\psi_{\mathrm{2D}}\) with respect to \(V_{\mathrm{G}}\), namely \(\beta\), is given by:
\begin{equation}
\label{eqn-def-beta}
\beta = \frac{\mathrm{d} \psi_{\mathrm{2D}}}{\mathrm{d} V_{\mathrm{G}}} 
      = \frac{\frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}}{\frac{1}{C_{\mathrm{d}}} + \frac{1}{C_{\mathrm{2D}}}}
\end{equation}
The index \(\beta\) here quantifies the tunability of the contact angle change
by \(V_{\mathrm{G}}\). Accordingly, a high degree of \(\beta\) can be attained by increasing both \(C_{\mathrm{d}}\) and
\(C_{\mathrm{2D}}\), thereby introducing the dependence on the choice of 2D material.

Here we demonstrate such phenomenon by considering an electrowetting
setup comprised of a thin, high-\(k\) dielectric layer (2 nm HfO\(_{\text{2}}\)
layer with the relative permittivity \(\epsilon_{\mathrm{d}}=24.0\))
underlying a layer of monolayer 2D material (Figure
\ref{fig:dcos-all-2D}(a)).  As addressed earlier, due to the fact that
\((\Delta \cos \theta)^{\mathrm{Orien}}\) of graphene-water system may
not be readily applied to other 2D materials, we compare the
calculated \((\Delta \cos \theta)^{\mathrm{EDL}}\) as a function of
\(V_{\mathrm{G}}\), by using the DFT-calculated \(C_{\mathrm{2D}} -
\sigma_{\mathrm{2D}}\) relations for a variety of 2D materials, as
shown in \ref{fig:dcos-all-2D}(b). Clearly, as a consequence of the high
quantum capacitance, the wettability of the 2D semiconductors
considered is more tunable compared to that for the 2D semimetals. In
other words, in a 2D semiconductor, the value required to reach a
certain is lowered. We predict that the contact angle change \(\Delta \cos \theta\) for
the 2D semiconductors (e.g. TMDCs) can reach up to 0.22\textasciitilde{}0.25 within the range of
\(V_{\mathrm{G}}\) considered, corresponding to an interfacial tension change of 15\textasciitilde{}18 \(\mathrm{mJ} \cdot \mathrm{m}^{-2}\). The analysis presented here suggests that the manipulation of a liquid
droplet on a layer of 2D material doped by an electric displacement field may be feasible
\cite{Mugele_2005,Hayes_2003_nature_EWOD}.

An interesting implication is that, together with the recent
development in engineering 2D materials’ wetting translucency
\cite{raj_wettability_2013,rafiee_wetting_2012,shih_breakdown_2012,shih_wetting_2013},
in principle, upon doping, a 2D material becomes less “transparent” (or more screening)
to both van der Waals and Coulombic interactions exerted from the underlying
substrate, due to enhanced liquid-2D material interactions. That is to say, the doping level can be
another control variable to modulate the molecular packing and
epitaxial behavior on a 2D material-coated surface, which may bring
new technology opportunity for a variety of applications.

\section{Conclusion}
\label{sec:org333a0f3}

In conclusion, we present a multiscale theoretical framework
concerning the wettability of doped 2D materials, by considering: (i)
the change of 2D materials surface energy, (ii) the molecular
reorientation of liquid molecules adjacent to the interface, and (iii)
the electrical double layer formed in the liquid phase. Taking
graphene as an example, we show that the Coulombic interaction
dominates the change of liquid-2D material interfacial tension, at
both atomistic and continuum length scales.The latter two effects were
found to be the major mechanisms responsible for the water contact angle
change at the 2D material interfaces upon doping.

The doping-induced reorientation of liquid molecules at the
graphene-liquid interface is revealed by MD simulations, which also
allow decoupling of vdW and Coulombic energy contributions to the
interfacial tension in the absence of electrolytes in liquid. It is
found that the interfacial energy change is dominated by the Coulombic
interactions. Our results also reveal an asymmetric change of
graphene-water interfacial energy upon doping, such that slightly
n-doped graphene can reorient interfacial water molecules to minimize
electrostatic attractions, and therefore, slightly increase the
interfacial energy.

On the hand, the EDL effect is calculated by a continuum model. Our
analysis suggests that the reorientation effect is more predominant
than the EDL effect, On the graphene-water interface, we predict that
the combined reorientation and EDL effects can induce a significant
change of the interfacial tension \(\Delta
\gamma_{\mathrm{2D-L}}\), up to -15\textasciitilde{}-20 \(\mathrm{mJ}\cdot
\mathrm{m}^{-2}\), at the doping level of \(\pm 4 \times 10^{13}\ e\cdot
\mathrm{cm}^{-2}\). By adding the fitting parameter concerning the
defect density, our theoretical framework can nicely describe the
experimentally observed doping-induced contact angle change. Finally,
based on the DFT-calculated quantum capacitances (QCs) for a variety
of 2D materials, we predict that the wettability of 2D semiconductors
(e.g., TMDCs) is more tunable under an electric displacement field,
compared with 2D semimetals (e.g. graphene) due to their high quantum
capacitances. Our findings reveal a complete picture for the
modulation of the molecular interactions between liquid and a 2D monolayer upon
doping. The multiscale theoretical
framework proposed here is expected to shed
light on the surface science of 2D materials, 
as well as to provide a quantitative estimation for the wettability
of the doped 2D materials. We hope that the development for the 2D
materials-based functional surfaces in liquid manipulation, energy
harvesting, and molecular epitaxy will be facilitated by the
fundamental principles and theoretical insights presented here.

\bibliography{ref}

\newpage{}
\section{Figures}
\label{sec:org0a7d2b4}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.95\linewidth]{../img/dgamma-sigma.pdf}
\caption{\label{fig:dgamma-sigma}
\(\Delta \gamma_{\mathrm{2D}}\) as a function of \(\sigma_{\mathrm{2D}}\) for selected 2D materials: graphene, silicene, germanene, phosphorene, MoS\(_{2}\), MoSe\(_{2}\), MoTe\(_{2}\), WS\(_{2}\), WSe\(_{2}\) and WTe\(_{2}\)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.95\linewidth]{../img/scheme-methods.png}
\caption{\label{fig:scheme-method}
Scheme of the multiscale approach for modeling the doping-induced wettability tuning of 2D materials.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{../img/fig-pot-dens.pdf}
\caption{\label{fig:MD-res}
Geometry of the periodic MD simulation box for (a) water molecules only and (b) water-graphene systems.(c) Change of total adhesion energy \(\Delta\Phi\), and its partial contributionsfrom Lennard-Jones interaction \(\Delta\Phi_{\mathrm{LJ}}\) and Coulombic interaction \(\Delta\Phi_{\mathrm{CL}}\), as a function of \(\sigma_{\mathrm{2D}}\). 10\(^{\text{-3}}\) \$e\$/atom corresponds to a charge density of \(3.80 \times 10^{12}\ e \cdot \mathrm{cm}^{-2}\). (d) Local density \(\rho_{\mathrm{L}}\) (e) local charge density \(\delta_{\mathrm{L}}\) of water molecules as a function of distance \(z\) from graphene surface.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\linewidth]{../img/2d-ph-dependency+MD.pdf}
\caption{\label{fig:res-EDL}
(a) Scheme of the interface between the 2D material and the aqueous phase. (b) \((\Delta\cos\theta)^{\mathrm{EDL}}\) as a function of \(\sigma_{\mathrm{2D}}\) with varied solute concentrations. The concentration \(c_{0}\) varies from \(10^{0}\) to \(10^{-7}\) mol\(\cdot\mathrm{L}^{-1}\) (c) Overall change of contact angle \((\Delta \cos \theta)^{\mathrm{Orien+EDL}}\) combining the orientation and EDL effects, with varied solute concentrations as in (b).}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.95\linewidth]{../img/plot-fitting.pdf}
\caption{\label{fig:f-nc-exp}
Schemes of the void region in (a) substrated doped 2D material and (b) electrostatic gated 2D material.(c) Fitted experimental data of \(\Delta\cos\theta\) as a function of \(\sigma_{\mathrm{2D}}\). The electrowetting data are extracted from Ref. \citenum{hong_mechanism_2016}; the ESEM data are extracted from Ref. \citenum{ashraf_doping-induced_2016}.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.65\linewidth]{../img/dcos-all-2D.pdf}
\caption{\label{fig:dcos-all-2D}
(a) Schematic illustration of the 2D-material-based electrowetting device, where the 2D material is electostatically doped. (b) \((\Delta\cos\theta)^{\mathrm{EDL}}\) as a function of \(V_{\mathrm{G}}\) for selected 2D materials.}
\end{figure}
\end{document}