### 2018-10-05 submission/Langmuir/20171017/latex_files/writing/manuscript_revised.tex

% Created 2017-08-07 Mon 15:10
% Intended LaTeX compiler: pdflatex
\documentclass[journal=langd5,manuscript=article,email=true,hyperref=true,keywords=true]{achemso}
\usepackage{graphicx}
\setkeys{acs}{maxauthors = 0}   %no truncation
\usepackage{float}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{fontspec}
\usepackage[section]{placeins}  %enable section number
\SectionNumbersOn
\author{Tian Tian}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\author{Shangchao Lin}
\email{slin@eng.fsu.edu.}
\affiliation{Department of Mechanical Engineering, Materials Science and Engineering Program, FAMU-FSU College of Engineering, Florida State University, Tallahassee, Florida 32310, United States}
\author{Siyu Li}
\affiliation{Key Laboratory of Energy Thermal Conversion and Control of Ministry of Education, School of Energy and Environment, Southeast University, Nanjing, Jiangsu 210096, China}
\author{Lingling Zhao}
\affiliation{Key Laboratory of Energy Thermal Conversion and Control of Ministry of Education, School of Energy and Environment, Southeast University, Nanjing, Jiangsu 210096, China}
\author{Elton J. G. Santos}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\author{Chih-Jen Shih}
\email{chih-jen.shih@chem.ethz.ch}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\keywords{two-dimensional materials, doping, wettability, multiscale analysis, molecular dynamics simulation, electrical double layer}
\date{}
\title{Doping-Driven Wettability of Two-Dimensional Materials: a Multiscale Theory}
\begin{document}

\begin{tocentry}
\includegraphics[width=8.6cm]{../img/TOC_Langmuir.pdf}
\end{tocentry}

\newpage{}
\begin{abstract}
  Engineering molecular interactions at two-dimensional (2D) materials
  interfaces enables new technological opportunities in functional
  surfaces and molecular epitaxy.  Understanding the wettability of 2D
  materials represents the crucial first step towards quantifying the
  interplay between the interfacial forces and electric potential of
  2D materials interfaces.  Here we develop the first theoretical
  framework to model the wettability of the doped 2D materials by
  properly bridging the multiscale physical phenomena at the 2D
  interfaces, including: (i) the change of 2D materials surface energy
  (atomistic scale, several \AA), (ii) the molecular reorientation of
  liquid molecules adjacent to the interface (molecular scale,
  \(10^{0} \sim 10^{1}\) nm), and (iii) the electrical double layer
  (EDL) formed in the liquid phase (mesoscopic scales,
  \(10^{0} \sim 10^{4}\) nm).  The latter two effects are found to be
  the major mechanisms responsible for the contact angle change upon
  doping, while the surface energy change of a pure 2D material has no
  net effect on the wetting property.  When the doping level is
  electrostatically tuned, we demonstrate that 2D materials with high
  quantum capacitances (e.g., transition metal dichalcogenides, TMDCs)
  possess a wider range of tunability in the interfacial tension,
  under the same applied gate voltage.  Furthermore, practical
  considerations such as defects and airborne contamination are also
  quantitatively discussed.  Our analysis implies that the doping
  level can be another variable to modulate the wettability at 2D
  materials interfaces, as well as the molecular packing behavior on a
  2D material-coated surface, essentially facilitating the interfacial
  engineering of 2D materials.
\end{abstract}
% \textbf{Keywords}
% two-dimensional materials, doping, wettability, multiscale analysis, molecular dynamics simulation, electrical double layer
 \maketitle

\newpage{}

\section{Introduction}
\label{sec:org6e779b8}

Two-dimensional (2D) materials, the crystalline films with a thickness
of one atom or one molecule, combine optical transparency and the
quantum-confined properties
\cite{Novoselov_2005,Mas_Ballest__2011,Novoselov_2016}.  Recent progress
in the scalable production of large-area 2D materials
\cite{Boott_2015,Li_2016} has paved the way for the applications of
functional surfaces and ultrathin membranes
\cite{Prasai_2012,Feng_2016,Rafiee_2010,Yin_2014,Surwade_2015,Hern_ndez_2013}.
The concept of van der Waals (vdW) epitaxy has been further extended
to grow molecular heterostructures on 2D material-coated surfaces for
a variety of applications
\cite{Shi_2012_epitaxy_graphene,Kim_2017_remote_epitaxy,Novoselov_2016}. The
realization of these technologies essentially requires precise control
over the molecular interactions at 2D material interfaces, which
motivates fundamental research of their interfacial properties, in
particular the wettability.  An accurate determination of the
interfacial tension, nevertheless, remains highly disputed
\cite{taherian2013what,Kozbial_2015,Parobek_2015,Govind_Rajan_2016}. Early
literature on understanding the surface science of 2D materials has
highlighted the effects of the airborne contaminants
\cite{li_effect_2013,Xu_2013_withwhat,kozbial_study_2014,Kozbial_2015,Chow_2015}
and the underlying substrate
\cite{raj_wettability_2013,rafiee_wetting_2012,shih_breakdown_2012,shih_wetting_2013},
since the length scale for the van der Waals (vdW) radius is
comparable to the thickness of a monolayer. On the other hand, a
subtle but important fact is that the 2D semimetals (e.g, graphene and
silicene) and 2D semiconductors (e.g., transition metal
dichalcogenides (TMDCs)) possess low density of states (DOS) around
the intrinsic Fermi level (\(E_{\mathrm{F}}\)), such that the effect of
doping, either induced by the surroundings
\cite{Chen_2013,Varchon_2007,Giovannetti_2008}, or by the electrostatic
gating \cite{Das_2008,Perera_2013}, also comes into play. Recent advance
in the \textit{ab initio} calculations and scanning tunneling
microscopy (STM) studies on the adsorbed molecules on graphene
\cite{Muruganathan_2015,Huttmann_2015} has provided some clues to the
doping-dependent vdW interactions. Subsequently, the doping-induced
change in the wettability of graphene has been observed in the
contact-angle experiments
\cite{hong_mechanism_2016,goniszewski_correlation_2016,ashraf_doping-induced_2016},
and the molecular dynamics (MD) simulations
\cite{ostrowski_tunable_2014,ren_interfacial_2015,Taherian_2015,daub_electrowetting_2007}. In
this respect, however, most reports described the doping-dependent
wettability with the basic Young-Lippmann equation (YLE)
\cite{Lippmann_1875}. The interplay between the orientation of
liquid molecules at the interface \cite{Shen_2006} and the electronic
structure of the 2D materials is often ignored. In order to address
the discrepancies among the literature, a general and complete
theoretical picture that bridges the gap between different length
scales, is clearly required.

In this report, we propose the first theoretical framework to model
the change of interfacial tension between liquid and a sheet of
monolayer 2D material. Multiscale physical phenomena are considered,
as shown in Figure \ref{fig:scheme-method}. At the atomistic scale
(several \AA{}), we formulate the dependence of surface energy of a 2D
material on the doping density using the quantum capacitance (QC)
calculated by the hybrid density functional theory (DFT) at the level
of Heyd-Scuseria-Ernzerhof functionals (HSE06). Next, the
surface-charge-induced reorientation of liquid molecules adjacent to
the interface is associated with an N-body system, which is resolved
by molecular dynamics (MD) simulations, allowing quantification of the
interfacial tension in the absence of electrolytes in liquid at the
molecular scale (\(10^{0} \sim 10^{1}\) nm). The effect of the electrical
double layer (EDL) induced by the electrostatic interactions between
the 2D material surface charges and the ionic species in liquid, is
then addressed at the continuum level (\(10^{0} \sim 10^{4}\) nm). Practical
considerations, such as the defect density and the surface
contamination are also taken into account to provide a comprehensive
understanding of the phenomena. Finally, we examine and validate our
theory by comparing with the contact angle changes reported in the
electrowetting and the substrate-induced doping experiments.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.95\linewidth]{../img/scheme-methods.pdf}
\caption{\label{fig:scheme-method} Scheme of the proposed multiscale
  approach for modeling the doping-induced wettability tuning of 2D
  materials. Multiscale phenomena are considered (right side, from top
  to bottom): (i) change of 2D material surface energy
  \(\Delta \gamma_{\mathrm{2D}}\) caused by quantum capacitance, at
  the atomistic scale (several \AA{}). (ii) change of interfacial
  energy by reorientation of interfacial liquid molecules
  \(\Delta \gamma_{\mathrm{2D-L}} ^{\mathrm{Orien}}\), at molecular
  scale (10\(^{\text{0}}\) \textasciitilde{} 10\(^{\text{1}}\)
  nm). (iii) change of interfacial energy as a result of electric
  double layer \(\Delta \gamma_{\mathrm{2D-L}} ^{\mathrm{EDL}}\),
  described by a continuum model (10\(^{\text{0}}\) \textasciitilde{}
  10\(^{\text{4}}\) nm)}
\end{figure}

\section{Results and Discussion}
\label{sec:orga11172d}

\subsection{Surface Energy of Doped 2D Materials}
\label{sec:org8df08f3}

Consider a liquid (\emph{L}) droplet sitting on a flat, monolayer 2D
material (\emph{2D}) supported by a solid substrate, following the Young’s
equation, the change of the equilibrium contact angle \(\theta\) upon
doping, is given by
\begin{equation}
\label{eqn-def-Young-Delta-theta}
\gamma_{\mathrm{L}} \Delta \cos\theta = \Delta \gamma_{\mathrm{2D}}
                                 - \Delta \gamma_{\mathrm{2D-L}}
\end{equation}
where \(\gamma_{\mathrm{L}}\) and \(\gamma_{\mathrm{2D}}\) are the surface
tensions of liquid and the 2D material considered, respectively,
\(\gamma_{\mathrm{2D-L}}\) is the interfacial tension between the liquid
and the contacting 2D material. We define \(\Delta \gamma = \gamma -
\gamma_{0}\) for all the surface (interfacial) tensions considered and
\(\Delta \cos \theta = \cos \theta - \cos \theta_{0}\), where the
subscript 0 corresponds to that in the case of intrinsic 2D material,
with the doping density per unit area, \(\sigma_{\mathrm{2D}} = 0\). In
the theoretical analysis presented here, we aim to model \(\Delta \cos
\theta\) as a function of \(\sigma_{\mathrm{2D}}\). Note that under the
assumptions that (i) the change of \(E_{\mathrm{F}}\) in the doped 2D
materials does not result in the interfacial electron transfer,
namely, the electrochemical reactions, (ii) the doping effect does not
change the surface energy of the underlying substrate, and (iii) the
vdW and electrostatic interactions are perfectly additive and
pairwise, one can decouple the effect of the underlying
substrate. Therefore the debate of the “wetting transparency”
\cite{rafiee_wetting_2012,shih_wetting_2013}, does not affect our
analysis of \(\Delta \cos \theta\), which is independent of \(\gamma_{0}\)
or \(\theta_{0}\).

First, we model the dependence of \(\gamma_{\mathrm{2D}}\) on the doping
level. Consider a closed system with constant pressure and volume
containing a sheet of free-standing 2D material, following the Euler
homogeneous function theorem of thermodynamics, the total internal
energy of the system \(U\), is given by \(U = TS + \mu_{\mathrm{2D}} N +
\gamma_{\mathrm{2D}} A + \psi_{\mathrm{2D}} q\)
\cite{bard_electrochemical_1980}, where \(T\) is the temperature, \(S\) is
the entropy, \(\mu_{\mathrm{2D}}\) is the chemical potential of the 2D
material per unit lattice, \(N\) is the number of unit lattices, \(A\) is
the area of the 2D material, and \(\psi_{\mathrm{2D}}\) and \(q\) are the
electric potential and total charge in the 2D material,
respectively. At constant \(T\), combining with the first law of
thermodynamics and the differential form of \(U\), it follows:
\begin{equation}
\label{eqn-dgamma-dpsi}
\mathrm{d} \gamma_{\mathrm{2D}} = -\frac{q}{A} \mathrm{d} \psi_{\mathrm{2D}}
                                = -\sigma_{\mathrm{2D}} \mathrm{d} \psi_{\mathrm{2D}}
\end{equation}
relating the surface tension change of a free-standing 2D material as
a function of its doping density \(\sigma_{\mathrm{2D}}\). After bringing an amount of charge
\(q\) from infinity to a charge-neutral 2D material in the
aforementioned system, the surface energy change is therefore given
by:
\begin{equation}
\label{eqn-delta-gamma-sigma-free-2D}
\Delta \gamma_{\mathrm{2D}} = - \int_{0}^{\psi_{\mathrm{2D}}} \sigma_{\mathrm{2D}} \mathrm{d}\psi'
                            = - \int_{0}^{\sigma_{\mathrm{2D}}} \sigma' \left( \frac{1}{C_{\mathrm{2D}}}\right) \mathrm{d} \sigma'
\end{equation}
where \(C_{\mathrm{2D}} = g(E_{\mathrm{F}}) e^{2}\) is the quantum
capacitance of the 2D material \cite{davies_two-dimensional_1997,Das_Sarma_2011},
\(g(E_{\mathrm{F}})\) is the DOS as a function of \(E_{\mathrm{F}}\),
\(\psi_{\mathrm{2D}} = -(E_{\mathrm{F}} - E_{\mathrm{F,0}})/e\) and
\(E_{\mathrm{F,0}}\) corresponds to the Fermi level of the 2D material
at the charge neutral point (CNP). Accordingly, eq
\ref{eqn-delta-gamma-sigma-free-2D} provides a simple relation which
depicts the surface tension change of a 2D material at the
quantum-mechanical level. We have calculated the DOS as a function of
\(E_{\mathrm{F}}\) for a variety of 2D materials using the density
functional theory \cite{tian_multiscale_2016}. Note that we take into
account a fractional component of the exact exchange from the
Hartree-Fock (HF) theory hybridized with the DFT exchange-correlation
functional at the level of the HSE06 hybrid functional. Therefore, any
limitation of the exchange and correlation functional utilized in the
chemical description of the energy levels can be improved. The
following 2D materials are considered: TMDC monolayers (MX\(_{\text{2}}\), M =
Mo, W and X = S, Se, Te), silicene, germanene, phosphorene (monolayer
black phosphorus), and graphene. The doping density in a 2D material
is calculated by integrating the DOS from its intrinsic Fermi level,
i.e., \(\sigma_{\mathrm{2D}} = \int_{E_{\mathrm{F,0}}}^{E_{\mathrm{F}}}
g(E')e \mathrm{d}E'\) \cite{john_quantum_2004}.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75\linewidth]{../img/dgamma-sigma.pdf}
\caption{\label{fig:dgamma-sigma} Atomistic scale:
  \(\Delta \gamma_{\mathrm{2D}}\) as a function of
  \(\sigma_{\mathrm{2D}}\) for selected 2D materials: graphene,
  silicene, germanene, phosphorene, MoS\(_{2}\), MoSe\(_{2}\),
  MoTe\(_{2}\), WS\(_{2}\), WSe\(_{2}\) and WTe\(_{2}\). The 2D
  semimetals (graphene, silicene and germanene) with low high quantum
  capcitances show larger decrease of \(\gamma_{\mathrm{2D}}\) upon
  doping, compared with the 2D semiconductors (TMDCs, phosphorene).}
\end{figure}

Figure \ref{fig:dgamma-sigma} presents the calculated
\(\Delta \gamma_{\mathrm{2D}}\) as a function of
\(\sigma_{\mathrm{2D}}\) for the 2D materials considered
here. Clearly, the doping of 2D materials reduces their surface
energy, or based on the classical definition, the work required to
separate two stacked monolayers is lowered. Among the 2D materials at
the same doping level, we find that graphene shows the highest degree
of surface energy decrease, up to -16 mJ\(\cdot \mathrm{m}^{-2}\) , or
\textasciitilde{}20\% reduction of its intrinsic surface tension
\cite{shih_wetting_2013}, at \(\sigma_{\mathrm{2D}}\) =
\(\pm 4\times10^{13}\ e\cdot \mathrm{cm}^{-2}\).  A clear trend is
that the surface energy decrease is more significant in the 2D
semimetals (e.g. graphene, silicene, and germanene) than that in the
2D semiconductors (e.g. TMDCs). This reflects the fact that the
effective mass of carriers in the 2D semiconductors is much higher
than that in the 2D semimetals \cite{davies_two-dimensional_1997},
thereby resulting in high DOS, as well as a high \(C_{\mathrm{2D}}\)
(see eq \ref{eqn-delta-gamma-sigma-free-2D}). This concept also
explains why the surface energy decrease for silicene and germanene
are lower than that for graphene \cite{Yan_2013}.  To our knowledge,
the doping-induced surface energy change in 2D materials has never
been investigated experimentally, which may be of interest for future
study, inspired by the recent advances in direct measurement of 2D
material surface energy \cite{van_Engers_2017_direct_surf_gr}. Note
that for modeling the macroscopic wetting behavior on a doped 2D
material interface, the reorientation and EDL effects also need to be
taken into account.  As will be discussed later, the doped-induced
change of \(\gamma_{\mathrm{2D}}\) does not imply a reduced
wettability, because the quantum capacitance effect also reduces the
interfacial tension, \(\gamma_{\mathrm{2D-L}}\), which yields zero net
effect on wettability.

\subsection{Reorientation of Liquid Molecules}
\label{sec:orgc4a9a93}

Next, we discuss the interactions between 2D materials and liquid. In
a doped 2D material, the delocalized carriers are confined in the 2D
plane. Therefore, following the spirit of the mean-field theory, we
treat it as a continuously, uniformly charged surface. Since these
charges are either generated by interacting with the underlying
substrate, or electrostatically induced by gating, the
electroneutrality still holds before in contact with liquid. The
surface charges result in two consequences that may change
\(\gamma_{\mathrm{2D-L}}\), including (i) the reorientation of adjacent
liquid molecules \cite{ostrowski_tunable_2014} and (ii) the formation of
the electric double layer (EDL) at the liquid-solid interface, known
as the electrowetting effect
\cite{Lippmann_1908,mugele_electrowetting:_2005}. Under the assumption
that the vdW and electrostatic (Coulombic) interactions are additive,
which compose the interfacial tension, we propose that the
interfacial tension change is given by:

\begin{equation}
\label{eqn-delta-gamma-decompose}
\Delta \gamma_{\mathrm{2D-L}} = \Delta \gamma_{\mathrm{2D-L}}^{\mathrm{Orien}}
                              + \Delta \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}}
\end{equation}
where \(\Delta \gamma_{\mathrm{2D-L}}^{\mathrm{Orien}}\) and \(\Delta
\gamma_{\mathrm{2D-L}}^{\mathrm{EDL}}\) correspond to the contributions
from the reorientation and the EDL effects, respectively. We also
assume that \(\Delta \gamma_{\mathrm{2D-L}}^{\mathrm{Orien}}\) is independent of
the electrolyte concentration, since the concentration of liquid
molecules is typically orders-of-magnitude higher.

First we discuss the orientation effect. Understanding the
reorientation effect involves positioning and sampling the collective,
time-averaged motion of liquid molecules near the interface, which is
a standard molecular dynamics (MD) problem. Note that even with the
state-of-the-art MD algorithms, it remains challenging to accommodate
the calculations for the EDL, in which the length scale of electric
field can be larger than one micrometer in diluted electrolyte
solutions, e.g., pure water with self-ionized H\(_{\text{3}}\)O\(^{\text{+}}\) and OH\(^{\text{-}}\)
ions. Here we consider the graphene-water interface as a model
system. All MD simulations were carried out using the GROMACS 4.5
software package \cite{Hess_2008}. Monolayer graphene was modeled as an
infinite rigid sheet in the x-y plane. The carbon atoms of graphene
were treated as uncharged Lennard-Jones (LJ) spheres with \(\sigma\) =
0.34 nm and \(\epsilon\) = 0.223 kJ/mol \cite{Cheng_1990}, using the
force-field parameters reported by Tummala and Striolo
\cite{Tummala_2008}. The doping effect is included by assigning an equal
amount of charge \(\sigma_{\mathrm{2D}} / \rho_{\mathrm{G}}\), where
\(\rho_{\mathrm{G}}\) is the surface density of carbon atoms, to each
carbon atom. The \(\sigma_{\mathrm{2D}}\) range considered here
approximately correspond to the partial atomic charge from −0.012 to
0.012 \(e/\mathrm{atom}\). Water molecules were modeled using the SPC/E
model \cite{Berendsen_1987} with bond lengths and angles of water
molecules constrained using the SETTLE algorithm
\cite{Miyamoto_1992}. Lennard-Jones interactions were treated with a
cutoff distance of 1 nm, with those between different atoms calculated
using the standard geometric averaging rule. Long-range electrostatic
interactions were treated using the particle mesh Ewald (PME)
summation method \cite{Darden_1993,Essmann_1995} with a short-range
cutoff distance of 1 nm. The velocity-rescaled Berendsen thermostat
was implemented to maintain a constant system temperature of 298.15 K
\cite{Bussi_2007}. All simulations were carried out under the NVT
ensemble, and the equations of motion of water molecules were
integrated over a range of 20 ns with 10\(^{\text{7}}\) steps.

In order to precisely determine the interfacial interactions using MD
simulations, instead of the commonly used model that compared the
nanoscale contact angle by placing a nanodroplet onto a sheet of
suspended 2D material
\cite{ostrowski_tunable_2014,daub_electrowetting_2007,ren_interfacial_2015,Taherian_2015},
we simulate the difference of the total potential energy, \(E\),
between two separate systems that contain (i) only water molecules
with two surfaces exposing to vacuum (L), and (ii) the same amount of
water molecules with one surface in contact with graphene (placed at z
= 0) and the other surface exposing to vacuum (GL) (see Figure
\ref{fig:MD-res}(a) and \ref{fig:MD-res}(b)). Note that the current
model can only reveal the static and macroscopic wetting behavior,
while the effects of dynamic properties or surface roughness, as
discussed by some recent MD-based nano-electrowetting studies
\cite{Yuan_2010_PF,Zhao_2015_pillar}, are not considered here.
Periodic boundary conditions are used in all three directions of the
simulation boxes in both systems. Additionally, a vacuum layer of 3 nm
thick along the \emph{z}-axis is placed to separate the periodic
images of the graphene-water system. The total energy in both systems
can be formulated as:
\(E_{\mathrm{L}} = \mu_{\mathrm{L}}n_{\mathrm{L}} +
2\gamma_{\mathrm{L}}S\) and
\(E_{\mathrm{GL}}=\mu_{\mathrm{L}}n_{\mathrm{L}}+(\gamma_{\mathrm{L}}
+ \gamma_{\mathrm{2D-L}} + \gamma_{\mathrm{2D}})S\), respectively,
where \(\mu_{\mathrm{L}}\) is the chemical potential per water
molecule in the bulk phase, \(n_{\mathrm{L}}\) is the number of liquid
molecules in the simulation box, and \(S\) is the area of the
\emph{xy}-plane.  A simulation box with a sufficiently large length of
21 nm in the \emph{z}-direction is used, with a 18-nm thick block of
water molecules, to minimize the effect of the long-range
electrostatic interaction between the charged graphene and the water
molecules at the water-vacuum interface, by ensuring that the
time-averaged dipole moment for the water molecules at the
water-vacuum interface approaches zero (see Supporting Information
Figure S1).  One can show that
\(E_{\mathrm{GL}} - E_{\mathrm{L}} = (\gamma_{\mathrm{2D-L}} +
\gamma_{\mathrm{2D}} - \gamma_{L})S = (\Phi + 2
\gamma_{\mathrm{2D}})S\), where \(\Phi\) is the interfacial energy,
which is defined as
\(\Phi = \gamma_{\mathrm{2D-L}} - \gamma_{\mathrm{2D}} -
\gamma_{\mathrm{L}}\), and combining with eq
\ref{eqn-def-Young-Delta-theta}, the change of interfacial energy
\(\Delta \Phi\) can be formulated as:
\begin{equation}
\label{eqn-Delta-Phi-Delta-cos}
\Delta \Phi = \Delta (E_{\mathrm{GL}} - E_{\mathrm{L}})/S - 2\Delta \gamma_{\mathrm{2D}} = -\gamma_{\mathrm{L}} (\Delta \cos \theta)^{\mathrm{Orien}}
\end{equation}
where \((\Delta\cos \theta)^{\mathrm{Orien}}\) corresponds to the
contact angle change due to the reorientation effect. Note that in the
current MD simulation setup, the surface tension of the 2D material
remains unchanged (i.e.  \(\Delta \gamma_{\mathrm{2D}}=0\)), since the
Fermi level change with respect to \(\sigma_{\mathrm{2D}}\) (see eq
\ref{eqn-delta-gamma-sigma-free-2D}) is not included in the algorithm.
When analyzing the reorientation effect on the MD level, we adopt the
water surface tension value of
\(\gamma_{\mathrm{L}}=63.6\ \mathrm{mJ}\cdot \mathrm{m}^{-2}\) in the
SPC/E model \cite{Vega_2007}. Note that the value of
\(\gamma_{\mathrm{L}}\) only affects the results of
\(\Delta \cos \theta\) while not the interaction potential \(\Phi\), as revealed by eq
\ref{eqn-Delta-Phi-Delta-cos}. Moreover, even though the surface
tension from SPC/E model is ~12.6\% lower than the real value (72.8
mJ\(\cdot\)m\(^2\)), such minor difference may not affect our
conclusion. Accordingly, \(\Delta \Phi\) follows
\(\Delta \Phi = \Delta (E_{\mathrm{GL}} - E_{\mathrm{L}})/S = \Delta
\Phi_{\mathrm{LJ}} + \Delta \Phi_{\mathrm{Coul}}\), where
\(\Delta \Phi_{\mathrm{LJ}}\) and \(\Delta \Phi_{\mathrm{Coul}}\)
correspond to the change for the vdW and Coulombic interaction
potentials, respectively.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{../img/fig-pot-dens.pdf}
\caption{\label{fig:MD-res} Molecular scale: geometry of the periodic
  MD simulation box for (a) water molecules only and (b)
  water-graphene systems, with \textasciitilde{} 27000 water molecules
  per unit cell. (c) Change of total adhesion energy \(\Delta\Phi\),
  and its partial contributionsfrom Lennard-Jones
  \(\Delta\Phi_{\mathrm{LJ}}\) and Coulombic
  \(\Delta\Phi_{\mathrm{Coul}}\) interactions, as a function of
  \(\sigma_{\mathrm{2D}}\). The filled regions indicate the error
  margin of calculation. \(10^{-3}\) \textit{e}/atom corresponds to a
  charge density of \(3.80 \times 10^{12}\ e \cdot
  \mathrm{cm}^{-2}\). (d) and (e) are the local density
  \(\rho_{\mathrm{L}}\) and local charge density
  \(\delta_{\mathrm{L}}\) of water molecules as a function of distance
  \(z\) from graphene surface, respectively.}
\end{figure}

Figure \ref{fig:MD-res}(c) shows the calculated \(\Delta \Phi\), and
its contributions from the Lennard-Jones (\(\Delta \Phi_{\mathrm{LJ}}\))
and Coulombic interactions (\(\Delta \Phi_{\mathrm{Coul}}\)), as
functions of \(\sigma_{\mathrm{2D}}\). We note that the change of
adhesion energy in the doped graphene system is dominated by the
Coulombic interaction. When the doping level of graphene is \(\pm 4
\times 10^{13}\ e\cdot \mathrm{cm}^{-2}\), the Coulombic interaction
causes a decrease in the adhesion energy of \textasciitilde{} -15 mJ\(\cdot
\mathrm{m}^{-2}\), while the vdW interaction, on the other hand, causes
a slight increase in the adhesion energy by less than 5 mJ\(\cdot
\mathrm{m}^{-2}\). In other words, concerning the reorientation of
water molecules at large doping levels, the Coulombic interaction
favors the decrease of \(\gamma_{\mathrm{2D-L}}\), and thus the contact
angle \(\theta\), while the vdW interaction slightly increases
\(\gamma_{\mathrm{2D-L}}\) and \(\theta\). We shall note that due to the
highly polar nature of water molecules, the average inter-molecular
equilibrium distance in the absence of external electric field
(\(r_{\mathrm{O-O}}=2.75\ \mathrm{\AA}\)) is shorter than the
Lennard-Jones equilibrium distance (\(r_{\mathrm{O-O}}=3.16\
\mathrm{\AA}\)) \cite{Mark_2001}. Increasing the doping density of the
graphene sheet essentially enhances the Coulombic attractions between
graphene and water molecules, further reducing the inter-molecular
equilibrium distance, as seen in Figure S2(a), the distance between
the first water layer and graphene (calculated from maximal
\(\rho_{\mathrm{L}}\) from Figure \ref{fig:MD-res}(d)) slightly decreases
in doped graphene systems, compared with the charge-neutral system,
which eventually leads to a slight increase in \(\Delta
\Phi_{\mathrm{LJ}}\).  The plot of \(\Delta \Phi\) as a function of
\(\sigma_{\mathrm{2D}}\) in Figure \ref{fig:MD-res}(c) shows an apparent
asymmetric behavior. This can be further revealed from the
\emph{z}-dependent local molecular density \(\rho_{\mathrm{L}}\) and charge
density \(\delta_{\mathrm{L}}\) profiles of the water molecules, as
shown in Figure \ref{fig:MD-res}(d) and \ref{fig:MD-res}(e),
respectively. The molecular density of the first water layer adjacent
to graphene increases when graphene is p-doped
(\(\sigma_{\mathrm{2D}}=0.012\ e\cdot \mathrm{cm}^{-2}\), or
equivalently 4.56 \textit{e}/atom) and decreases when graphene is
n-doped (\(\sigma_{\mathrm{2D}}=-0.012\ e\cdot \mathrm{cm}^{-2}\), or
equivalently -4.56 \textit{e}/atom) compared with the case of
charge-neutral graphene, indicating the polarity-dependent adsorption
of water molecules on graphene: lighter H atoms prefer to adsorb onto
n-doped graphene electrostatically. Similar trend can also be observed
in the \(\delta_{\mathrm{L}}\) (Figure \ref{fig:MD-res}(e))and water
dipole orientation \(\cos \mu\) (Figure S1) profiles, both
suggesting that water orientation depends strongly on
\(\sigma_{\mathrm{2D}}\): more negative \(\sigma_{\mathrm{2D}}\) tends
to orient the water dipoles more towards the 2D interface, reflected
by the more positive charge densities due to first-layer H atoms, and
the more negative \(\cos \mu\) values. Interestingly, the \(\Delta
\Phi_{\mathrm{Coul}}\) exhibits small positive contribution for slightly
n-doped graphene, when \(\sigma_{\mathrm{2D}} \approx -2\times10^{13}\
e\cdot \mathrm{cm}^{-2}\) or equivalently -0.76 \textit{e}/atom,
indicating that the reorientation of interfacial water molecules by
small negative charges on graphene helps minimizing the Coulombic
interactions. The trend observed in hydrogen-bond densities (see
Figure S3) also matches those observed in \(\rho_{\mathrm{L}}\),
\(\delta_{\mathrm{L}}\) and \(\cos \mu\) profiles.

Finally, we believe that the asymmetric behavior of \(\Delta
\Phi_{\mathrm{Coul}}\) with respect to \(\sigma_{\mathrm{2D}}\) is due to
the nonlinear decrease in \(\delta_{\mathrm{L}}\) of the first water
layer, when graphene becomes more n-doped (see Figure S2(d)).  As a
general trend, \(\delta_{\mathrm{L}}\) of the first water layer
increases when \(\sigma_{\mathrm{2D}}\) becomes more negative, due to
the attraction of H atoms to the graphene surface. However we observe
a decrease of \(\delta_{\mathrm{L}}\) when \(\sigma_{\mathrm{2D}}\)
ranges from -0.001 to -0.006 \(e \cdot \mathrm{cm}^{-2}\) (see Figure
S2(d)), which could be a combined result of the graphene-water
Coulombic interactions and intermolecular hydrogen bonding. We also
notice a nonlinear behavior in the \(\rho_{\mathrm{L}}\) of the first
water layer: when \(\sigma_{\mathrm{2D}}\) becomes more negative, the
interatomic distance \(r_{\mathrm{C-O}}\) between graphene C and water O
atoms will increase due to Coulombic repulsion.  Note that the
increased Coulombic repulsion energy for C−O pairs from more negative
\(\sigma_{\mathrm{2D}}\) values is balanced by the increased vdW
attractions between C−O pairs when \(r_{\mathrm{C-O}}\) increases, which
eventually prevents the reduction of the first water layer density at
an equilibrium C-O distance, as seen at \(\sigma_{\mathrm{2D}} < -0.01\
e \cdot \mathrm{cm}^{-2}\) (Figure S2(b)). Here we propose a simple
capacitor-like model, considering the contributions from
\(\delta_{\mathrm{L}}\) of the first water layer of the first water
layer and \(\sigma_{\mathrm{2D}}\) to the interfacial Coulombic energy
\(\Phi_{\mathrm{Coul}}^{\mathrm{int}}\) (see the discussion in
Supporting Information and Figure S4). We show that in the n-doped graphene system
the charge density of the first water layer has a major contribution
to \(\Delta \Phi_{\mathrm{Coul}}\). The nonlinear behavior of
\(\delta_{\mathrm{L}}\) at low n-doping level is responsible for the
shift of maximum of the \(\Delta
\Phi_{\mathrm{Coul}}-\sigma_{\mathrm{2D}}\) curve. The analysis further
indicates the importance of interfacial water configuration in
determining the change of the interaction potentials.

Although the MD simulation results show that the Coulombic
interactions between the doped graphene and water molecules dominate
the reorientation effect, such conclusion may not be readily applied
to other types of 2D materials, as the intensity of vdW interactions
and surface charge polarization can vary greatly. First, the
microscopic vdW and Coulombic interactions on 2D materials depend on
the lattice constant and atomic density. In addition, for homoatomic
2D materials such as graphene, silicene, germanene and phosphorene,
their surface charge polarization is negligible compared to
heteroatomic 2D materials, for instance TMDC. It has been shown that,
for TMDCs, the value of \(\gamma_{\mathrm{2D-L}}\) is already largely
determined by the Coulombic interactions in the charge-neutral systems
\cite{Govind_Rajan_2016,Chow_2015}. Therefore we anticipate that
case-by-case studies and DFT calculations which deal with the
post-doping charge redistribution, might be required, as shown by
several previous studies
\cite{Sresht_2015,Govind_Rajan_2016,Sresht_2017}.

\subsection{Electrical Double Layer Effect}
\label{sec:org492fdee}

The EDL effect that decreases the interfacial tension has been
extensively studied in the context of the electrowetting phenomena
\cite{Mugele_2005}, in which the interfacial tension is reduced due to
the adsorption of ionic species at the solid-liquid interface. Here we
extend the concept to the 2D material-liquid systems. Consider a
closed system containing an interface formed between liquid and a
sheet of free-standing 2D material, by combining with the Gibbs
adsorption theory \cite{bard_electrochemical_1980}, one can obtain:
\begin{equation}
\label{eq:2}
\mathrm{d} \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}} = -\sigma_{\mathrm{2D}} \mathrm{d} \psi_{\mathrm{2D}}
                                                   -\sum_{\mathrm{i}} \Gamma_{\mathrm{i}} \mathrm{d}\mu_{\mathrm{i}}
\end{equation}
where \(\Gamma_{\mathrm{i}}\) and \(\mu_{\mathrm{i}}\) are the interfacial
excess and the chemical potential of the ionic specie \(i\) at the
interface, respectively. Relative to that in the bulk phase
(\(\psi=0\)), the interfacial chemical potential of a charged solute
\(\mu_{\mathrm{i}}\), is given by
\(\mu_{\mathrm{i}}=z_{\mathrm{i}}e\psi_{\mathrm{2D}}\), where
\(z_{\mathrm{i}}\) is the valency of component \(i\), under the assumption
that the interfacial electric potential is equal to
\(\psi_{\mathrm{2D}}\). As with the approach leading to eq
\ref{eqn-delta-gamma-sigma-free-2D}, it follows:
\begin{equation}
\label{eqn-Delta-gamma-2DL-EDL-full}
\Delta \gamma_{\mathrm{2D-L}}^{\mathrm{EDL}} = 
-\int_{0}^{\sigma_{\mathrm{2D}}} \sigma' \left(\frac{1}{C_{\mathrm{2D}}}\right) \mathrm{d}\sigma'
-\int_{0}^{\sigma_{\mathrm{L}}} \sigma' \left(\frac{1}{C_{\mathrm{EDL}}}\right) \mathrm{d}\sigma'
\end{equation}
where \(\sigma_{\mathrm{L}}\) is the interfacial charge density per unit
area in the liquid phase, and \(C_{\mathrm{EDL}}\) is the EDL
capacitance. Combining eq \ref{eqn-Delta-gamma-2DL-EDL-full} with eqs
\ref{eqn-def-Young-Delta-theta} and \ref{eqn-delta-gamma-sigma-free-2D},
the contact angle change due to the EDL effect, \((\Delta \cos
\theta)^{\mathrm{EDL}}\) is given by:
\begin{equation}
\label{eqn-Delta-cos-EDL}
(\Delta \cos \theta)^{\mathrm{EDL}} = -\frac{1}{\gamma_{\mathrm{L}}}
                                      \int_{0}^{\sigma_{\mathrm{L}}} \sigma' 
                                      \left(\frac{1}{C_{\mathrm{EDL}}}\right) \mathrm{d}\sigma'
\end{equation}
In other words, eqs \ref{eqn-delta-gamma-sigma-free-2D} and
\ref{eqn-Delta-gamma-2DL-EDL-full} suggest that \(\sigma_{\mathrm{2D}}\)
decreases both \(\gamma_{\mathrm{2D}}\) and \(\gamma_{\mathrm{2D-L}}\) to
the same extent, i.e. \(\int_{0}^{\sigma_{\mathrm{2D}}}
\sigma'/C_{\mathrm{2D}} \mathrm{d}\sigma'\), such that the contribution
of quantum capacitance cancels out in the YLE.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\linewidth]{../img/2d-ph-dependency+MD.pdf}
\caption{\label{fig:res-EDL} The continuum model: (a) Scheme of the
  interface between the 2D material and the aqueous phase. (b)
  \((\Delta\cos\theta)^{\mathrm{EDL}}\) as a function of
  \(\sigma_{\mathrm{2D}}\) with varied solute concentrations. The
  concentration \(c_{0}\) varies from \(10^{-1}\) to \(10^{-7}\)
  mol\(\cdot\mathrm{L}^{-1}\) (c) Overall change of contact angle
  \((\Delta \cos \theta)^{\mathrm{Orien+EDL}}\) combining the
  orientation and EDL effects, with varied solute concentrations as in
  (b).}
\end{figure}

For a \(z:z\) electrolyte solution, with the molar concentration \(c_{0}\)
and the electrolyte valence \(z\), we describe the electric potential
profile in liquid with the one-dimensional Gouy-Chapman-Stern model
\cite{Stern_1924_theory}, as schematically shown in Figure
\ref{fig:res-EDL}(a).  The EDL consists of a Helmholtz layer with the
permittivity of liquid \(\epsilon_{\mathrm{L}}\) and thickness
\(d_{\mathrm{H}}\), next to the Gouy-Chapman (GC) diffuse layer where
the electric potential profile follows the Poisson Boltzmann equation
\cite{bard_electrochemical_1980}. The \(C_{\mathrm{EDL}}\) of the
Gouy-Chapman-Stern model is equivalent to two capacitors in series,
namely,
\begin{equation}
\label{eqn-GCS-capacitance-series}
\frac{1}{C_{\mathrm{EDL}}} = \frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}
\end{equation}
where \(C_{\mathrm{H}}=\epsilon_{\mathrm{L}}/d_{\mathrm{H}}\) is the
capacitance of the Helmholtz double layer and
\(C_{\mathrm{GC}}=\sqrt{\frac{2z^{2}e^{2}\epsilon_{\mathrm{L}}c_{0}N_{\mathrm{A}}}{k_{\mathrm{B}}T}}
\cosh(\frac{ze\psi_{\mathrm{GC}}}{2k_{\mathrm{B}}T})\) is the
differential capacitance of the GC layer, with \(\psi_{\mathrm{GC}}\)
corresponding to the electric potential at the GC interface. Consider
a Gaussian enclosure including the 2D material layer and the EDL, due
to a zero net electric field at the surface of the Gaussian enclosure,
the electroneutrality of the system \cite{bard_electrochemical_1980}
suggests:
\begin{equation}
\label{eqn-neutrality}
\sigma_{\mathrm{2D}} + \sigma_{\mathrm{L}} = 0
\end{equation}
And therefore, the electric potentials corresponding to the surface of
the 2D material and the outer Helmholtz plane, \(\psi_{\mathrm{2D}}\)
and \(\psi_{\mathrm{GC}}\), are given by:
\begin{equation}
\label{eqn-psi-GC}
\begin{aligned}
  \psi_{\mathrm{GC}} &= -\frac{2k_{\mathrm{B}}T}{ze} 
                         \sinh^{-1}\left(
                         \frac{-\sigma_{\mathrm{2D}}}{\sqrt{8c_{0}N_{\mathrm{A}}\epsilon_{\mathrm{L}}k_{\mathrm{B}}T}}
                            \right) \\
  \psi_{\mathrm{2D}} &= \psi_{\mathrm{GC}} + \sigma_{\mathrm{2D}}\frac{d_{\mathrm{H}}}{\epsilon_{\mathrm{L}}}
\end{aligned}
\end{equation}
where \(N_{\mathrm{A}}\) is the Avogadro constant and \(k_{\mathrm{B}}\)
is the Boltzmann constant. Accordingly, the final form
of \((\Delta \cos \theta)^{\mathrm{EDL}}\) is given by:
\begin{equation}
\label{eqn-Delta-cos-EDL-final}
\begin{aligned}
(\Delta \cos \theta)^{\mathrm{EDL}} &= \frac{\sigma_{\mathrm{2D}}^{2}}{2\gamma_{\mathrm{L}}
                                       C_{\mathrm{H}}}\frac{1}{\gamma_{\mathrm{L}}}
                                       \sqrt{\frac{32k_{\mathrm{B}}^{3}T^{3} \epsilon_{\mathrm{L}} c_{0} 
                                       N_{\mathrm{A}}}{z^{2}e^{2}}} 
                                       \left[\cosh(\frac{z e \psi_{\mathrm{GC}}}{2k_{\mathrm{B}} T}) -1 \right]\\
                                    &= \frac{\Delta \sigma_{\mathrm{2D}}^{2}}{2\gamma_{\mathrm{L}}}
                                       \left[ \frac{1}{C_{\mathrm{H}}} + \frac{1}{(C_{\mathrm{DH}} + 
                                       C_{\mathrm{GC}})/2}\right]
\end{aligned}
\end{equation}
where \(C_{\mathrm{DH}}=\epsilon_{\mathrm{L}}/\lambda_{\mathrm{D}}\) is
the differential capacitance of the diffuse layer using the
Debye-Hückel approximation \cite{bard_electrochemical_1980}, and
\(\lambda_{\mathrm{D}}=\sqrt{\frac{\epsilon_{\mathrm{L}}k_{\mathrm{B}}T}{2z^{2}e^{2}c_{0}N_{\mathrm{A}}}}\)
is the Debye screening length. Compared with the classical YLE,
\(\Delta \cos \theta =
-\frac{\sigma^{2}}{2\gamma_{\mathrm{L}}C_{\mathrm{YL}}}\), where
\(C_{\mathrm{YL}}\) is the equivalent capacitance in YLE, one can show that \(C_{\mathrm{YL}}\)
is given by:
\begin{equation}
\label{eqn-CYL-model}
\frac{1}{C_{\mathrm{YL}}} = \frac{1}{C_{\mathrm{H}}} + \frac{1}{(C_{\mathrm{DH}} + C_{\mathrm{GC}})/2}
\end{equation}

The above analysis leads to two important observations: (i) If a thin
contamination layer (usually composed of charge-neutral hydrocarbon
species \cite{li_effect_2013}) exists between the Helmholtz layer and
the 2D material surface, the electroneutrality at the interface (eq
\ref{eqn-neutrality}) still holds, such that \((\Delta \cos
\theta)^{\mathrm{EDL}}\) remains the same (see eq
\ref{eqn-Delta-cos-EDL-final}). In other words, despite the recent
debate about the role of the airborne adsorbates in the contact angle
measurement on 2D materials
\cite{li_effect_2013,Kozbial_2015,Xu_2013_withwhat,Chow_2015}, we expect
that the experimental characterization of the doping-induced contact
angle change is more robust and reproducible. (ii) As shown in eqs
\ref{eqn-GCS-capacitance-series} and \ref{eqn-Delta-gamma-2DL-EDL-full},
the equivalent capacitance in the YLE, \(C_{\mathrm{YL}}\), is not
identical to the total capacitance in the EDL, \(C_{\mathrm{EDL}}\). We
notice that the literature in this field (e.g. Refs.
\citenum{ostrowski_tunable_2014,daub_electrowetting_2007,goniszewski_correlation_2016,ashraf_doping-induced_2016})
often simply assumed that \(C_{\mathrm{YL}}=C_{\mathrm{DH}}\) in the
YLE, which significantly overestimates \((\Delta \cos
\theta)^{\mathrm{EDL}}\), when \(\psi_{\mathrm{2D}} \gg
k_{\mathrm{B}}T/e\) \cite{israelachvili03_inter}, as can be found in heavily-doped 2D materials.

Consider the interface formed between graphene and an (1:1) aqueous
solution, using the parameters of \(d_{\mathrm{H}}=3\) Å
\cite{mcclendon_thickness_1927},
% and
% \(\gamma_{\mathrm{L}}=72.8\ \mathrm{mJ}\cdot \mathrm{m}^{-2}\) at
% \(T=298\) K,
the calculated \((\Delta \cos \theta)^{\mathrm{EDL}}\) as a function
of \(\sigma_{\mathrm{2D}}\) is shown in Figure \ref{fig:res-EDL}(b).
Moderate electrolyte concentrations \(c_{0}\) between \(10^{-1}\) to
\(10^{-7}\) mol\(\cdot\)L\(^{-1}\) are considered, since the corresponding Debye
screening length will be larger than the effective radius of vdW
interactions (to keep the classical continuum-based EDL model valid), under which decoupling the reorientation and EDL effects is
meaningful. Recent experimental and theoretical studies reveal that
the screening lengths of higher concentrations may increase instead of
exponential decay in classical models
\cite{Smith_2016_screening_length,Lee_2017_screening_length}, which
indicates that the addition of reorientation and EDL effects may still be
valid for higher concentrations, provided an accurate description of
the potential profile in the EDL. Nevertheless to keep our multiscale
model simple and generally applicable, we limit the electrolyte
concentration to the range mentioned above. Note that
\(c_{0}=10^{-7}\) mol\(\cdot \mathrm{L}^{-1}\) corresponds to the pure
water system. We find that the contact angle change is stronger in a
more diluted electrolyte solution, resulting from an increase in Debye
screening length \(\lambda_{\mathrm{D}}\), which in turn decreases
both \(C_{\mathrm{GC}}\) and \(C_{\mathrm{DH}}\). This is distinct
from the that in the typical electrowetting experiment on a dielectric
layer, in which \(C_{\mathrm{YL}}\) is governed by the capacitance of
the dielectric layer, so the contact angle change is almost
independent of the electrolyte concentration
\cite{Mugele_2005}. Moreover, within the range of
\(\sigma_{\mathrm{2D}}\) considered here (\(-4\times10^{13}\)
\textasciitilde{} \(4\times10^{13}\ e\cdot \mathrm{cm}^{-2}\)), the
calculated \((\Delta \cos \theta)^{\mathrm{EDL}}\) is always less than
0.06.  In combination with the
\(\Delta \gamma^{\mathrm{Orien}}_{\mathrm{2D-L}}\) (equivalent to
\(\Delta \Phi\), see previous section) calculated by the MD
simulations, the total change of contact angle
\((\Delta \cos \theta)^{\mathrm{Orien + EDL}}\) due to the doping
effect is shown in Figure \ref{fig:res-EDL}(c). We separately fit the
\(\Delta \Phi_{\mathrm{LJ}}\) and \(\Delta \Phi_{\mathrm{Coul}}\) from
the reorientation effect as functions of \(\sigma_{\mathrm{2D}}\) in
n- and p-doped regimes, respectively (see Figure S5), and use the
fitting results to construct the \(\Delta \Phi-\sigma_{\mathrm{2D}}\)
relation. Comparing the interactions at different length scales, we
find that the reorientation effect is more predominant than the EDL
effect at moderate electrolyte concentrations, while the shift of the
Fermi level in 2D materials has no net effect on the wetting, due to the
cancellation effect explained previously. Therefore the overall change
of contact angle is less dependent on the solute concentration,
compared with Figure \ref{fig:res-EDL}(b).

Together with the reorientation effect, the calculated \(\Delta \cos
\theta\) change only reaches 0.06 \textasciitilde{} 0.08, corresponding to a maximum
contact angle decrease of 3.5 \textasciitilde{} 4.5\(^{\circ}\), for
\(\sigma_{\mathrm{2D}}=\pm1\times10^{13}\ e\cdot \mathrm{cm}^{-2}\),
which is is a typical range for both the electrostatic gating
\cite{hong_mechanism_2016} and substrate-induced doping
\cite{ashraf_doping-induced_2016}. Clearly, the predicted contact angle
change upon doping is still lower than those observed
experimentally. In the next section, we discuss the practical
considerations that may influence a 2D material-coated surface.

\subsection{Practical Considerations}
\label{sec:orgb997597}

In practice, in order to
carry out the contact angle measurement, a large-area sheet of 2D
material is required to cover the surface. Therefore, most of the
experimental findings are based on the chemical-vapor-deposited (CVD)
samples, which are inherently polycrystalline with grain boundaries
and point defects \cite{Banhart_2011_defect}.  In addition, the transfer
of the ultrathin films onto a solid surface often results in
nanometer- to micrometer- scale tears and wrinkles. Although our
analysis has suggested that the charge-neutral polymer residues and
airborne contaminations do not influence \((\Delta
\theta)^{\mathrm{EDL}}\), here we show that the defect density plays an
important role in determining the wettability of doped 2D
materials.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.95\linewidth]{../img/plot-fitting.pdf}
\caption{\label{fig:f-nc-exp} Comparison between model and
  experimental data: schemes of the void region in (a) substrated
  doped 2D material and (b) electrostatic gated 2D material.(c) Fitted
  experimental data of \(\Delta\cos\theta\) as a function of
  \(\sigma_{\mathrm{2D}}\). The electrowetting data are extracted from
  Ref. \citenum{hong_mechanism_2016}; the ESEM data are extracted from
  Ref. \citenum{ashraf_doping-induced_2016}.}
\end{figure}

Figures \ref{fig:f-nc-exp}(a) and \ref{fig:f-nc-exp}(b) illustrate how
an electric field, either from the dopants on the substrate surface or
from the electrostatic gating, penetrate through a void in a 2D
material sheet and interact with the liquid phase directly. As a
result, the EDL is built up adjacent to the substrate surface, with
the surface excess \(\sigma_{\mathrm{v}}\) and the effective
capacitance \(C_{\mathrm{v}}\).  For molecular doping, the EDL charges
in the void regions are built up by the fixed charges on the substrate
surface, while in the electrostatic doping, the electric field from
the charged 2D material to the gate electrode through the void regions
is responsible for building up the EDL. For both cases, we assume that
\(\sigma_{v}\) is equivalent to \(\sigma_{\mathrm{2D}}\), due to
charge conservation. Together with the EDL and the reorientation
effects discussed earlier, the modified YLE considering a 2D material
with voids follows:
\begin{equation}
\label{eqn-def-Delta-cos-mixture}
\Delta \cos \theta = -f\frac{\sigma_{\mathrm{2D}}^{2}}{\gamma_{\mathrm{L}} C_{\mathrm{v}}} 
                     + (1-f)[(\Delta \cos \theta)^{\mathrm{Orien}} + (\Delta \cos \theta)^{\mathrm{EDL}}]
\end{equation}
where \(f\) is the void (defect) fraction in the 2D material. For the
electrostatic gating experiments (see Figure
\ref{fig:f-nc-exp}(a)), the 2D material quantum capacitor
and the dielectric  capacitor are connected in
series, so that the voltage applied between the gate electrode and
2D material, \(V_{\mathrm{G}}\), is given by \cite{tian_multiscale_2016}:
\begin{equation}
\label{eqn-VG-gating}
V_{\mathrm{G}} = \frac{\sigma_{\mathrm{2D}} - \sigma_{\mathrm{0}}}{C_{\mathrm{d}}}
                  + \int_{\sigma_{0}}^{\sigma_{\mathrm{2D}}} \frac{1}{C_{\mathrm{2D}}} \mathrm{d}\sigma'
\end{equation}
where \(C_{\mathrm{d}}\) is the capacitance of the dielectric layer, and
\(\sigma_{0}\) is the initial doping density of the 2D material,
corresponding to \(V_{\mathrm{G}}=0\).

Next, in order to examine the effect of incomplete 2D material
coverage, two independent sets of experimental results, which measure
the water contact angle on (i) substrate-doped graphene
\cite{ashraf_doping-induced_2016} and (ii) electrostatically-gated
graphene \cite{hong_mechanism_2016} are chosen to compare, with
\(C_{\mathrm{v}} = C_{\mathrm{DH}}\) and \(C_{\mathrm{v}} =
C_{\mathrm{d}}\) in eq \ref{eqn-def-Delta-cos-mixture}, respectively. The
parameter \(f\) and \(\sigma_{0}\) are determined by least-square fitting
the experimentally-observed \(\Delta \cos \theta\) with respect to
\(\sigma_{\mathrm{2D}}\) using eq \ref{eqn-def-Delta-cos-mixture}. Figure
\ref{fig:f-nc-exp}(c) compares the calculated \(\Delta \cos \theta\) as a
function of \(\sigma_{\mathrm{2D}}\), together with the two sets of
experimental data. In both cases considered, we observe a slight
degree of shift in the minima of the fitted curves, corresponding to
the CNP of graphene, or \(\sigma_{\mathrm{2D}} = 0\). In other words, we
observe \(\sigma_{0} \neq 0\) in both cases, which is typical for
CVD-grown graphene samples
\cite{Shih2015PartiallyScreened,goniszewski_correlation_2016}.  The
fitted values of \(f\) are reasonably small (3.6\% for the
electrostatically-gated graphene and 3.0\% for substrate-doped
graphene), clearly demonstrating that the contact angle change can be
greatly influenced by the defect density. We believe this explains the
discrepancy between the experimental observations and the multiscale
theoretical framework proposed here.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.65\linewidth]{../img/dcos-all-2D.pdf}
\caption{\label{fig:dcos-all-2D} Electrostatic tunability of 2D
  material wettability: (a) Schematic illustration of the
  2D-material-based electrowetting device, where the 2D material is
  electostatically doped. (b) \((\Delta\cos\theta)^{\mathrm{EDL}}\) as
  a function of \(V_{\mathrm{G}}\) for selected 2D materials.}
\end{figure}

Finally, we discuss the influence of 2D material choice under
electrostatic gating condition. The above analysis (eq
\ref{eqn-Delta-cos-EDL-final}) has clearly suggested that the contact
angle change effect \((\Delta \cos \theta)^{\mathrm{EDL}}\) only depends
on \(\sigma_{\mathrm{2D}}\), which can be controlled by an electric
displacement field, with the experimental setup shown in Figure
\ref{fig:dcos-all-2D}(a). Specifically,  eqs \ref{eqn-psi-GC} and
\ref{eqn-VG-gating} suggest:
\begin{equation}
\label{eqn-dVG-choice-2D}
\begin{aligned}
\mathrm{d} V_{\mathrm{G}} &= (\frac{1}{C_{\mathrm{d}}} + \frac{1}{C_{\mathrm{2D}}}) \mathrm{d} \sigma_{\mathrm{2D}} \\
\mathrm{d} \psi_{\mathrm{2D}} &= -(\frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}) \mathrm{d} \sigma_{\mathrm{L}}
\end{aligned}
\end{equation}
And since \(\sigma_{\mathrm{2D}} = -\sigma_{\mathrm{L}}\), the first
derivative of \(\psi_{\mathrm{2D}}\) with respect to \(V_{\mathrm{G}}\), namely \(\beta\), is given by:
\begin{equation}
\label{eqn-def-beta}
\beta = \frac{\mathrm{d} \psi_{\mathrm{2D}}}{\mathrm{d} V_{\mathrm{G}}} 
      = \frac{\frac{1}{C_{\mathrm{H}}} + \frac{1}{C_{\mathrm{GC}}}}{\frac{1}{C_{\mathrm{d}}} + \frac{1}{C_{\mathrm{2D}}}}
\end{equation}
The index \(\beta\) here quantifies the tunability of the contact angle change
by \(V_{\mathrm{G}}\). Accordingly, a high degree of \(\beta\) can be attained by increasing both \(C_{\mathrm{d}}\) and
\(C_{\mathrm{2D}}\), thereby introducing the dependence on the choice of 2D material.

Here we demonstrate such phenomenon by considering an electrowetting
setup comprised of a thin, high-\(k\) dielectric layer (2 nm
HfO\(_{\text{2}}\) layer with the relative permittivity
\(\epsilon_{\mathrm{d}}=24.0\)) underlying a layer of monolayer 2D
material (Figure \ref{fig:dcos-all-2D}(a)).  As addressed earlier, due
to the fact that \((\Delta \cos \theta)^{\mathrm{Orien}}\) of
graphene-water system may not be readily applied to other 2D
materials, we compare the calculated
\((\Delta \cos \theta)^{\mathrm{EDL}}\) as a function of
\(V_{\mathrm{G}}\), by using the DFT-calculated
\(C_{\mathrm{2D}} - \sigma_{\mathrm{2D}}\) relations for a variety of
2D materials, as shown in \ref{fig:dcos-all-2D}(b). Clearly, as a
consequence of the high quantum capacitance, the wettability of the 2D
semiconductors considered is more tunable compared to that for the 2D
semimetals under same the applied gate voltage. In other words, in a 2D
semiconductor, the value required to reach a certain is lowered. We
predict that the contact angle change \(\Delta \cos \theta\) for the
2D semiconductors (e.g. TMDCs) can reach up to
0.22\textasciitilde{}0.25 within the range of \(V_{\mathrm{G}}\)
considered, corresponding to an interfacial tension change of
15\textasciitilde{}18 \(\mathrm{mJ} \cdot \mathrm{m}^{-2}\). The
analysis presented here suggests that the manipulation of a liquid
droplet on a layer of 2D material doped by an electric displacement
field may be feasible \cite{Mugele_2005,Hayes_2003_nature_EWOD}.

An interesting implication is that, together with the recent
development in engineering 2D materials’ wetting translucency
\cite{raj_wettability_2013,rafiee_wetting_2012,shih_breakdown_2012,shih_wetting_2013},
in principle, upon doping, a 2D material becomes less “transparent” (or more screening)
to both van der Waals and Coulombic interactions exerted from the underlying
substrate, due to enhanced liquid-2D material interactions. That is to say, the doping level can be
another control variable to modulate the molecular packing and
epitaxial behavior on a 2D material-coated surface, which may bring
new technological opportunity for a variety of applications.

\section{Summary and Conclusions}
\label{sec:org0fb53ec}

In conclusion, we present a multiscale theoretical framework
concerning the wettability of doped 2D materials, by considering: (i)
the change of 2D materials surface energy, (ii) the molecular
reorientation of liquid molecules adjacent to the interface, and (iii)
the electrical double layer formed in the liquid phase. Taking
graphene as an example, we show that the Coulombic interaction
dominates the change of liquid-2D material interfacial tension, at
both molecular and mesoscopic length scales. The latter two effects were
found to be the major mechanisms responsible for the water contact angle
change at the 2D material interfaces upon doping.

The doping-induced reorientation of liquid molecules at the
graphene-liquid interface is revealed by MD simulations, which also
allow decoupling of vdW and Coulombic energy contributions to the
interfacial tension in the absence of electrolytes in liquid. It is
found that the interfacial energy change is dominated by the Coulombic
interactions. Our results also reveal an asymmetric change of
graphene-water interfacial energy upon doping, such that slightly
n-doped graphene can reorient interfacial water molecules to minimize
electrostatic attractions, and therefore, slightly increase the
interfacial energy. Such asymmetric interfacial energy change arises
from the dipolar nature of water, which implies that the polarities of
liquid and 2D material both determine the symmetry of
doping-driven wettability: the change of interfacial energy upon
doping between nonpolar liquid molecules and homoatomic 2D material is
more symmetric than that between polar liquid and polar 2D
material. We expect that in the latter case, the symmetry of
interfacial energy change would highly depend on the delicate balance
between their degrees of polarization.

On the other hand, the EDL effect is calculated by a continuum model. Our
analysis suggests that the reorientation effect is more predominant
than the EDL effect. On the graphene-water interface, we predict that
the combined reorientation and EDL effects can induce a significant
change of the interfacial tension \(\Delta
\gamma_{\mathrm{2D-L}}\), up to -15 \(\mathrm{mJ}\cdot
\mathrm{m}^{-2}\), at the doping level of \(\pm 4 \times 10^{13}\ e\cdot
\mathrm{cm}^{-2}\). By adding the fitting parameter concerning the
defect density, our theoretical framework can nicely describe the
experimentally observed doping-induced contact angle change. Finally,
based on the DFT-calculated quantum capacitances (QCs) for a variety
of 2D materials, we predict that the wettability of 2D semiconductors
(e.g., TMDCs) is more tunable under an electric displacement field,
compared with 2D semimetals (e.g. graphene) due to their high quantum
capacitances. Our findings reveal a complete picture for the
modulation of the molecular interactions between liquid and a 2D monolayer upon
doping. The multiscale theoretical
framework proposed here is expected to shed
light on the surface science of 2D materials, 
as well as to provide a quantitative estimation for the wettability
of the doped 2D materials. We hope that the development for the 2D
materials-based functional surfaces in liquid manipulation, energy
harvesting, and molecular epitaxy will be facilitated by the
fundamental principles and theoretical insights presented here.

\begin{acknowledgement}
  T.T and C.-J. S. are grateful for the financial support from the ETH
  startup funding.  S.L. would like to acknowledge the startup funding
  from the Energy and Materials Initiative at the Florida State
  University. S.L. and L.Z. would like to acknowledge funding from the
  National Natural Science Foundation of China (Grant
  No. 51106027). E.J.G.S. acknowledges the use of computational
  resources from the UK national high performance computing service,
  ARCHER, for which access was obtained via the UKCP consortium and
  funded by EPSRC grant ref EP/K013564/1; and the Extreme Science and
  Engineering Discovery Environment (XSEDE), supported by NSF grants
  number TG-DMR120049 and TG-DMR150017. The Queen’s Fellow Award
  through the startup grant number M8407MPH is also acknowledged.
\end{acknowledgement}

\begin{suppinfo}
Additional results for the MD simulation and analysis of nonlinear
behavior of interfacial Coulombic interactions in charged
graphene-water system.
\end{suppinfo}

\bibliography{ref}

\end{document}